---
title: "Surg_onc Analysis"
output: html_document
date: "2025-04-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Surgical Oncology Analysis 

ensemble_FINAL_surgonc_May8 <- read.csv("~/Documents/R projects/tensorflow_probs3/ensemble_FINAL_surgonc_May8.csv", row.names=1)
ensemble_FINAL_surgonc = ensemble_FINAL_surgonc_May8
```{r}
ensemble_FINAL_surgonc = final_check_withtext

```

#Libraries
```{r}
library(data.table)
library(tidyr)
library(igraph)
library(dplyr)
library(tidygraph)
library(ggraph)
library(networkD3)
library(htmlwidgets)
library(shinydashboard)
library(shiny)
library(tableone)
library(flextable)
library(gtsummary)
library(ggpubr)
library(rworldmap)
library(scales)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(stringr)
library(RColorBrewer)
library(tibble)
library(viridis)
library(circlize)
```
```{r}

eligibilities <- read.table(file = "~/Documents/Research/Clinical Trials Study/March 2024 Clinical Trials Data /eligibilities.txt",
                   header = TRUE,
                   sep = "|",
                   na.strings = "",
                   colClasses = c("character", rep("character",13) ),  
                   comment.char = "",
                   quote = "\"",
                   fill = FALSE) 


facilities <- read.table(file = "~/Documents/Research/Clinical Trials Study/March 2024 Clinical Trials Data /facilities.txt",
                   header = TRUE,
                   sep = "|",
                   na.strings = "",
                   colClasses = c("character", rep("character",7) ),  
                   comment.char = "",
                   quote = "\"",
                   fill = FALSE) 

studies <- read.table(file = "~/Documents/Research/Clinical Trials Study/March 2024 Clinical Trials Data /studies.txt",
                   header = TRUE,
                   sep = "|",
                   na.strings = "",
                   colClasses = c("character", rep("character",6) ),  
                   comment.char = "",
                   quote = "\"",
                   fill = FALSE) 

countries <- read.table(file = "~/Documents/Research/Clinical Trials Study/March 2024 Clinical Trials Data /countries.txt",
                   header = TRUE,
                   sep = "|",
                   na.strings = "",
                   colClasses = c("integer", rep("character",3) ),  
                   comment.char = "",
                   quote = "\"",
                   fill = FALSE)  


interventions <- read.table(file = "~/Documents/Research/Clinical Trials Study/March 2024 Clinical Trials Data /interventions.txt",
                   header = TRUE,
                   sep = "|",
                   na.strings = "",
                   colClasses = c("integer", rep("character",4) ),  
                   comment.char = "",
                   quote = "\"",
                   fill = FALSE) 

conditions <- read.table(file = "~/Documents/Research/Clinical Trials Study/March 2024 Clinical Trials Data /conditions.txt",
                   header = TRUE,
                   sep = "|",
                   na.strings = "",
                   colClasses = c("integer", rep("character",3) ),  
                   comment.char = "",
                   quote = "\"",
                   fill = FALSE) 

sponsors <- read.table(file = "~/Documents/Research/Clinical Trials Study/March 2024 Clinical Trials Data /sponsors.txt",
                   header = TRUE,
                   sep = "|",
                   na.strings = "",
                   colClasses = c("integer", rep("character",4) ),  
                   comment.char = "",
                   quote = "\"",
                   fill = FALSE) 

calculated_values <- read.table(file = "~/Documents/Research/Clinical Trials Study/March 2024 Clinical Trials Data /calculated_values.txt",
                   header = TRUE,
                   sep = "|",
                   na.strings = "",
                   colClasses = c("integer", rep("character",18) ),  
                   comment.char = "",
                   quote = "\"",
                   fill = FALSE) 


browse_conditions <- read.table(file = "~/Documents/Research/Clinical Trials Study/March 2024 Clinical Trials Data /browse_conditions.txt",
                   header = TRUE,
                   sep = "|",
                   na.strings = "",
                   colClasses = c("integer", rep("character",4) ),  
                   comment.char = "",
                   quote = "\"",
                   fill = FALSE) 


designs <- read.table(file = "~/Documents/Research/Clinical Trials Study/March 2024 Clinical Trials Data /designs.txt",
                   header = TRUE,
                   sep = "|",
                   na.strings = "",
                   colClasses = c("integer", rep("character",13) ),  
                   comment.char = "",
                   quote = "\"",
                   fill = FALSE) 

drop_withdrawals <- read.table(file = "~/Documents/Research/Clinical Trials Study/March 2024 Clinical Trials Data /drop_withdrawals.txt",
                   header = TRUE,
                   sep = "|",
                   na.strings = "",
                   colClasses = c("integer", rep("character",9) ),  
                   comment.char = "",
                   quote = "\"",
                   fill = FALSE) 

```

#Countries analysis
```{r}

ensemble_countries_surgonc = merge(ensemble_FINAL_surgonc,countries, by="nct_id")
ensemble_countries_surgonc = dplyr::filter(ensemble_countries_surgonc, class=="yes")
ensemble_countries_surgonc = ensemble_countries_surgonc[!is.na(ensemble_countries_surgonc$name),]

```
#Rworldmap
```{r}

country_mapping = ensemble_countries_surgonc %>% distinct(nct_id, name) %>% count(name)

sPDF <- joinCountryData2Map(country_mapping, joinCode = "NAME", nameJoinColumn = "name", verbose = TRUE)
#Need to look at which countries do not map and fix their names using countrySynonyms 
#"Czechia" -> "Czech Republic"
#"Former Serbia and Montenegro" -> "Serbia and Montenegro"
#"Libyan Arab Jamahiriya" ->  "Libya"
#"Martinique" = "Martinique"
#"North Macedonia" = "Macedonia"

ensemble_countries_surgonc$name = str_replace_all(ensemble_countries_surgonc$name, c("Czechia" = "Czech Republic"))
ensemble_countries_surgonc$name = str_replace_all(ensemble_countries_surgonc$name, c("Former Serbia and Montenegro" = "Serbia"))
ensemble_countries_surgonc$name = str_replace_all(ensemble_countries_surgonc$name, c("Libyan Arab Jamahiriya" = "Libya"))
ensemble_countries_surgonc$name = str_replace_all(ensemble_countries_surgonc$name, c("Martinique" = "France"))
ensemble_countries_surgonc$name = str_replace_all(ensemble_countries_surgonc$name, c("North Macedonia" = "Macedonia"))

country_mapping = ensemble_countries_surgonc %>% distinct(nct_id, name) %>% count(name)

country_mapping_plot <- joinCountryData2Map(country_mapping, joinCode = "NAME", nameJoinColumn = "name", verbose = TRUE)

country_mapping = as.data.frame(country_mapping_plot@data)
compiled_country = country_mapping_plot@data
country_mapping = dplyr::select(country_mapping, c("ADMIN", "n", "name"))

world <- ne_countries(scale = "small", returnclass = "sf")

country_mapping_world <- merge(world, country_mapping, by.x="admin", by.y="ADMIN", all.x = FALSE, all.y = TRUE)

country_mapping_world = country_mapping_world[!is.na(country_mapping_world$name.y),]

cols = (brewer.pal(9, "YlOrRd"))
colfunc <- colorRampPalette(c(cols[1],cols[6], cols[9]), bias=2)
show_col(colfunc(100))
cols=colfunc(100)

world_map = world %>% dplyr::filter(!sovereignt=="Antarctica") %>%
ggplot() + 
  geom_sf(fill="white", linewidth=0.2) + 
  geom_sf(data = country_mapping_world,aes(fill=n)) +
  scale_fill_gradientn(
    colours = cols,
    name = "# Studies"
  ) +
  theme_classic() + 
  theme(legend.position = "bottom",
           legend.title = element_text(color = "black", size = 10),
           legend.text = element_text(color = "black", size = 10),
           legend.key.width = unit(2.0, 'cm'))

ggsave("world_map_surg_onc.svg", world_map, width=6, height=6)

```

#Plot of top # countries by study volume
```{r}
top20 <- country_mapping_world %>%
  st_drop_geometry() %>%
  select(name.y, n) %>%
  arrange(desc(n)) %>%
  slice_head(n = 20)


top20$name.y <- factor(top20$name.y, levels = rev(top20$name.y))

cols = (brewer.pal(9, "OrRd"))
colfunc <- colorRampPalette(c(cols[3],cols[7], cols[9]), bias=2)
show_col(colfunc(100))
cols=colfunc(100)

top20_plot = ggplot(top20, aes(x = name.y, y = n, fill = n)) +
  geom_col() +
  scale_fill_gradientn(
    colours = cols,
    name = "# Studies"
  ) +
  coord_flip() +
  ggtitle("Top 20 Countries") +
  xlab("Country") + 
     ylab("Number of Trials") +
  scale_y_continuous(limits = c(0,8200)) +
     theme_classic()  + 
     theme(
    panel.grid.major.y = element_line(color = "grey80", size = 0.4),
    panel.grid.major.x = element_line(color = "grey90", size = 0.4),
    panel.grid.minor = element_blank()
  ) +
     theme(axis.text.x = element_text(colour="black", size = 17)) +
     theme(axis.text.y = element_text(colour="black",size = 17)) + 
     theme(plot.title = element_text(colour="black", size=17,hjust = 0)) +
     theme(axis.title.x = element_text(colour="black", size =17)) +
     theme(axis.title.y = element_text(colour="black", size=17)) +
     theme(legend.position = "none")
top20_plot
ggsave("top20_plot.svg", top20_plot, width=5, height=6.5)

```


#Compile World bank data 
```{r}
worldbank_incomegroup <- read.csv("~/Documents/R projects/tensorflow_probs3/World Data/worldbank_incomegroup.csv", comment.char="#")
colnames(worldbank_incomegroup) <- gsub("\\.", "_", colnames(worldbank_incomegroup))

worldbank_population <- read.csv("~/Documents/R projects/tensorflow_probs3/World Data/worldbank_population.csv")
View(worldbank_population)
colnames(worldbank_population) <- gsub("^X", "pop_", colnames(worldbank_population))
colnames(worldbank_population) <- gsub("\\.", "_", colnames(worldbank_population))
worldbank_population = dplyr::select(worldbank_population, -c("Indicator_Name", "Indicator_Code"))

world_bank_GDP <- read.csv("~/Documents/R projects/tensorflow_probs3/World Data/world_bank_GDP.csv")
colnames(world_bank_GDP) <- gsub("^X", "gdp_", colnames(world_bank_GDP))
colnames(world_bank_GDP) <- gsub("\\.", "_", colnames(world_bank_GDP))
world_bank_GDP = dplyr::select(world_bank_GDP, -c("Country_Name", "Indicator_Name", "Indicator_Code"))

world_bank_compile = merge(worldbank_incomegroup,worldbank_population, by="Country_Code" )
world_bank_compile = merge(world_bank_compile,world_bank_GDP, by="Country_Code" )
world_bank_compile = dplyr::select(world_bank_compile, -"X")
```
#Compile health expeditures and health care system
```{r}
Health_expenditures <- openxlsx::read.xlsx("World Data/Health_expenditures.xlsx", sheet = "Data")
Health_expenditures = dplyr::select(Health_expenditures, "location", "code", "region", "year", "che_gdp", "hc71_usd")

# Filter the most recent year for each variable 
Health_expenditures_filtered <- Health_expenditures %>% 
  group_by(location) %>%                  # Group by 'Variable'
  filter(year == max(year)) %>%           # Keep only rows with the most recent year
  ungroup()  

```
#Compile health expeditures and health care system
```{r}
WHO_healthindicators <- openxlsx::read.xlsx("World Data/WHO_healthindicators.xlsx", sheet = "data")

WHO_healthindicators_filtered <- WHO_healthindicators %>% filter(IND_NAME %in% c("Density of medical doctors (per 10 000 population) ", "Life expectancy at birth (years)", "Total net official development assistance to medical research and basic health sectors per capita (US$), by recipient country ", "Population with household expenditures on health > 25% of total household expenditure or income (%)"))
                                                           
WHO_healthindicators_filtered <- WHO_healthindicators_filtered %>% dplyr::filter(is.na(DIM_1_CODE) | DIM_1_CODE == "SEX_BTSX")  %>% droplevels() 


# Filter the most recent year for each variable 
WHO_healthindicators_filtered <- WHO_healthindicators_filtered %>% 
  group_by(IND_NAME) %>%                  # Group by 'Variable'
  filter(DIM_TIME_YEAR == max(DIM_TIME_YEAR)) %>%           # Keep only rows with the most recent year
  ungroup()  

WHO_healthindicators_filtered = dplyr::select(WHO_healthindicators_filtered, -c("VALUE_COMMENTS", "IND_CODE", "DIM_1_CODE", "VALUE_STRING", "DIM_TIME_YEAR"))

WHO_healthindicators_filtered_final <- WHO_healthindicators_filtered %>%
  tidyr::pivot_wider(names_from = IND_NAME, values_from = VALUE_NUMERIC)

WHO_healthindicators_filtered_final
colnames(WHO_healthindicators_filtered_final) <- gsub(" ", "_", colnames(WHO_healthindicators_filtered_final))

```
#Health system 
```{r}
health_cluster <- openxlsx::read.xlsx("World Data/Table_cluster_vs_highest_value.xlsx", sheet = "2017_2")
health_cluster = dplyr::select(health_cluster, -c("Discrepancy", "GFA_D", "CHI_D", "OOP_D"))

colnames(health_cluster) <- str_replace_all(colnames(health_cluster), fixed("."), "_")

```


#Final compilation of global metrics
```{r}
global_compile = merge(world_bank_compile, WHO_healthindicators_filtered_final, by.x = "Country_Code", by.y = "DIM_GEO_CODE",all.x = TRUE, all.y = TRUE)

global_compile = merge(global_compile, Health_expenditures_filtered, by.x = "Country_Code", by.y = "code",all.x = TRUE, all.y = TRUE)

global_compile = merge(global_compile, health_cluster, by.x = "Country_Code", by.y = "Country_Code", all.x = TRUE, all.y = TRUE)

compiled_country = dplyr::select(compiled_country, c("ISO3", "n", "name" )) 

global_compile = merge(compiled_country, global_compile, by.x = "ISO3", by.y = "Country_Code", all.x = TRUE, all.y = TRUE)

global_compile = dplyr::select(global_compile, c("ISO3", "n", "name", "Region", "IncomeGroup", "pop_2000", "gdp_2023","pop_2020", "pop_2023", "Life_expectancy_at_birth_(years)", "Total_net_official_development_assistance_to_medical_research_and_basic_health_sectors_per_capita_(US$),_by_recipient_country_", "Density_of_medical_doctors_(per_10_000_population)_", "Population_with_household_expenditures_on_health_>_25%_of_total_household_expenditure_or_income_(%)" ,"year",  "che_gdp", "hc71_usd", "SHI_as_%_THE","Government_financing_as_%_of_THE","OOP_as_%_of_THE","VHI_as_%_of_THE","Predominance_defined_by_clustering", "Predominance_as_\"largest_HFS\""   ))

economy = dplyr::select(country_mapping_world, c("name.y", "economy"))

global_compile = merge(global_compile, economy, by.x = "name", by.y = "name.y", all.x = TRUE, all.y = TRUE)

global_compile <- global_compile %>%
  mutate(across(where(is.character), ~na_if(., "")))

global_compile = global_compile %>% filter(!is.na(name)) 
```

#Evaluation of Income group and Economy Type  with trial volume
```{r}
global_compile$economy = as.factor(global_compile$economy)

economy_plot = global_compile %>% filter(!is.na(economy)) %>% filter(!is.na(n)) %>% mutate(economy = forcats::fct_rev(economy)) %>%
  ggbarplot(., x = "economy", y = "n", fill = "economy", group = "economy", colour= "economy",
            palette = "Spectral", 
            add = "mean", 
            position = position_dodge(0.8),
            xlab = "Economy",
            ylab = "# Studies ",
            title = "Economy Type",
            panel.labs.font = list(size = 14)) +
  stat_summary(fun.data = mean_cl_boot, geom='errorbar', color='black', width=0.2, linewidth=0.75) +
  stat_compare_means(method = "kruskal.test", 
                     label.y = 1000, 
                     label.x = 2, 
                     size = 5) + 
  font("xlab", size = 13.5, color = "black") +
  font("ylab", size = 13.5, color = "black") +
  font("x.text", size = 14, color = "black") + 
  font("y.text", size = 14, color = "black") + 
  font("title", size = 13.5, color = "black") + coord_flip() +
  rremove("legend")

ggsave("economy_plot.png", economy_plot, width=5.5, height=3)
ggsave("economy_plot.svg", economy_plot, width=5.5, height=3)


#######
global_compile$IncomeGroup = as.factor(global_compile$IncomeGroup)

global_compile$IncomeGroup <- factor(global_compile$IncomeGroup, levels = c("High income", "Upper middle income", "Lower middle income", "Low income"))

incomegrp_plot = global_compile %>% filter(!is.na(IncomeGroup)) %>% filter(!is.na(n)) %>% mutate(IncomeGroup = forcats::fct_rev(IncomeGroup)) %>%
  ggbarplot(., x = "IncomeGroup", y = "n", fill = "IncomeGroup", group = "IncomeGroup", colour= "IncomeGroup",
            palette = "Spectral", 
            add = "mean", 
            position = position_dodge(0.8),
            xlab = "Income Group",
            ylab = "# Studies ",
            title = "Income Group",
            panel.labs.font = list(size = 14)) +
  stat_summary(fun.data = mean_cl_boot, geom='errorbar', color='black', width=0.2, linewidth=0.75) +
  stat_compare_means(method = "kruskal.test", 
                     label.x = 2, 
                     label.y = 300,
                     size=5) + 
  font("xlab", size = 13.5, color = "black") +
  font("ylab", size = 13.5, color = "black") +
  font("x.text", size = 14, color = "black") + 
  font("y.text", size = 14, color = "black") + 
  font("title", size = 13.5, color = "black") + coord_flip() +
  rremove("legend")

ggsave("incomegrp_plot.png", incomegrp_plot, width=5.5, height=3)
ggsave("incomegrp_plot.svg", incomegrp_plot, width=5.5, height=3)

```
#Health system spending correlations 
```{r}

names(global_compile)[names(global_compile) == 'SHI_as_%_THE'] <- 'SHI_perc_THE'
names(global_compile)[names(global_compile) == 'OOP_as_%_of_THE'] <- 'OOP_perc_THE'
names(global_compile)[names(global_compile) == 'VHI_as_%_of_THE'] <- 'VHI_perc_THE'
names(global_compile)[names(global_compile) == 'Government_financing_as_%_of_THE'] <- 'Gov_perc_THE'


```
#Health system type
```{r}
#Voluntary health insurance
#Government 
#Out Of Pocket
# (i.e., predominantly out-of-pocket, social health insurance (SHI) or government-financed),

shi_plot = global_compile  %>% filter(!is.na(SHI_perc_THE)) %>% filter(!is.na(n)) %>%
  ggplot(., aes(x=SHI_perc_THE, y=n)) +
     geom_point(size = 4.5,stroke = 1,colour="black",shape=21, fill= alpha("#38af30", 0.4)) + 
     geom_smooth(method = "lm", size = 3.25, color = "white", se = FALSE) +
     geom_smooth(method="lm", aes(colour=NULL, shape=NULL), size = 2.25, colour="#E41A1C", se =FALSE, alpha=0.8)+
     ggtitle("") +
     ylab("# Studies") +
     xlab("Social Health Insurance (%) ") +
     scale_x_continuous(limits = c(0,100)) +
     stat_cor(inherit.aes=FALSE,
                  method="pearson", 
                  data = global_compile,
                  aes(x = SHI_perc_THE, y =n),
                  cor.coef.name = "r",
                  r.digits = 2,
                  p.digits = 2,
                  size= 5,
                  ) +
     theme_classic()  + 
     theme(axis.text.x = element_text(colour="black", size = 16)) +
     theme(axis.text.y = element_text(colour="black",size = 16)) + 
     theme(plot.title = element_text(colour="black", size=16,hjust = 0)) +
     theme(axis.title.x = element_text(colour="black", size =16)) +
     theme(axis.title.y = element_text(colour="black", size=16)) +
     theme(strip.text = element_text(size=16)) +
     theme(legend.title = element_text(color = "black", size = 12),
                  legend.text = element_text(color = "black", size = 12),
                  legend.position = "none",
                  legend.direction="vertical",
                  legend.background = element_rect(size=0.5, linetype="solid", colour ="black"))


ggsave("shi_plot.png", shi_plot, width=3.5, height=3)
ggsave("shi_plot.svg", shi_plot, width=3.5, height=3)


vhi_plot = global_compile  %>% filter(!is.na(VHI_perc_THE)) %>% filter(!is.na(n)) %>%
  ggplot(., aes(x=VHI_perc_THE, y=n)) +
     geom_point(size = 4.5,stroke = 1,colour="black",shape=21, fill= alpha("#38af30", 0.4)) + 
  geom_smooth(method = "lm", size = 3.25, color = "white", se = FALSE) +
     geom_smooth(method="lm", aes(colour=NULL, shape=NULL), size = 2.25, colour="#E41A1C", se =FALSE, alpha=0.8)+
     ggtitle("") +
     ylab("# Studies") +
     xlab("Voluntary Insurance (%)") +
     scale_x_continuous(limits = c(0,100)) +
     stat_cor(inherit.aes=FALSE,
                  method="pearson", 
                  data = global_compile,
                  aes(x = VHI_perc_THE, y =n),
                  cor.coef.name = "r",
                  r.digits = 2,
                  p.digits = 2,
                  size= 5,
                  ) +
     theme_classic()  + 
     theme(axis.text.x = element_text(colour="black", size = 16)) +
     theme(axis.text.y = element_blank()) + 
     theme(plot.title = element_text(colour="black", size=16,hjust = 0)) +
     theme(axis.title.x = element_text(colour="black", size =16)) +
     theme(axis.title.y = element_blank()) +
     theme(strip.text = element_text(size=16)) +
     theme(legend.title = element_text(color = "black", size = 12),
                  legend.text = element_text(color = "black", size = 12),
                  legend.position = "none",
                  legend.direction="vertical",
                  legend.background = element_rect(size=0.5, linetype="solid", colour ="black"))


ggsave("vhi_plot.png", vhi_plot, width=3.5, height=3)
ggsave("vhi_plot.svg", vhi_plot, width=3.5, height=3)


oop_plot = global_compile  %>% filter(!is.na(OOP_perc_THE)) %>% filter(!is.na(n)) %>%
  ggplot(., aes(x=OOP_perc_THE, y=n)) +
     geom_point(size = 4.5,stroke = 1,colour="black",shape=21, fill= alpha("#38af30", 0.4)) + 
  geom_smooth(method = "lm", size = 3.25, color = "white", se = FALSE) +
     geom_smooth(method="lm", aes(colour=NULL, shape=NULL), size = 2.25, colour="#E41A1C", se =FALSE, alpha=0.8)+
     ggtitle("") +
     ylab("# Studies") +
     xlab("Out of Pocket (%)") +
     scale_x_continuous(limits = c(0,100)) +
     stat_cor(inherit.aes=FALSE,
                  method="pearson", 
                  data = global_compile,
                  aes(x = OOP_perc_THE, y =n),
                  cor.coef.name = "r",
                  r.digits = 2,
                  p.digits = 2,
                  size= 5,
                  ) +
     theme_classic()  + 
     theme(axis.text.x = element_text(colour="black", size = 16)) +
     theme(axis.text.y = element_blank()) + 
     theme(plot.title = element_text(colour="black", size=16,hjust = 0)) +
     theme(axis.title.x = element_text(colour="black", size =16)) +
     theme(axis.title.y = element_blank()) +
     theme(strip.text = element_text(size=16)) +
     theme(legend.title = element_text(color = "black", size = 12),
                  legend.text = element_text(color = "black", size = 12),
                  legend.position = "none",
                  legend.direction="vertical",
                  legend.background = element_rect(size=0.5, linetype="solid", colour ="black"))


ggsave("oop_plot.png", oop_plot, width=3.5, height=3)
ggsave("oop_plot.svg", oop_plot, width=3.5, height=3)


gov_plot = global_compile  %>% filter(!is.na(Gov_perc_THE)) %>% filter(!is.na(n)) %>%
  ggplot(., aes(x=Gov_perc_THE, y=n)) +
     geom_point(size = 4.5,stroke = 1,colour="black",shape=21, fill= alpha("#38af30", 0.4)) +
  geom_smooth(method = "lm", size = 3.25, color = "white", se = FALSE) +
     geom_smooth(method="lm", aes(colour=NULL, shape=NULL), size = 2.25, colour="#E41A1C", se =FALSE, alpha=0.8)+
     ggtitle("") +
     ylab("# Studies") +
     xlab("Government Funded (%)") +
     scale_x_continuous(limits = c(0,100)) +
     stat_cor(inherit.aes=FALSE,
                  method="pearson", 
                  data = global_compile,
                  aes(x = Gov_perc_THE, y =n),
                  cor.coef.name = "r",
                  r.digits = 2,
                  p.digits = 2,
                  size= 5,
                  ) +
     theme_classic()  + 
     theme(axis.text.x = element_text(colour="black", size = 16)) +
     theme(axis.text.y = element_blank()) + 
     theme(plot.title = element_text(colour="black", size=16,hjust = 0)) +
     theme(axis.title.x = element_text(colour="black", size =16)) +
     theme(axis.title.y = element_blank()) +
     theme(strip.text = element_text(size=16)) +
     theme(legend.title = element_text(color = "black", size = 12),
                  legend.text = element_text(color = "black", size = 12),
                  legend.position = "none",
                  legend.direction="vertical",
                  legend.background = element_rect(size=0.5, linetype="solid", colour ="black"))


ggsave("gov_plot.png", gov_plot, width=3.5, height=3)
ggsave("gov_plot.svg", gov_plot, width=3.5, height=3)

library(patchwork)

healthexpend_plots_patch =   shi_plot + gov_plot + vhi_plot + oop_plot + plot_layout(ncol = 4) 

ggsave("healthexpend_plots_patch.png", healthexpend_plots_patch, width=12, height=3)
ggsave("healthexpend_plots_patch.svg", healthexpend_plots_patch, width=12, height=3)

```

#Population graph 
```{r}
pop_plot = global_compile %>% filter(!is.na(n)) %>%
  ggplot(., aes(x=pop_2023, y=n)) +
     geom_smooth(method="lm", aes(colour=NULL, shape=NULL), size = 2.25, colour="grey60", se =FALSE)+
     geom_point(size = 4,stroke = 1,colour="black",shape=21, fill= alpha("#2b94db", 0.4)) + 
     ggtitle("Population vs Study (n)") +
     ylab("# Studies") +
     xlab("Population (Millions)") +
     stat_cor(inherit.aes=FALSE,
                  method="pearson", 
                  data = global_compile,
                  aes(x = pop_2023, y =n),
                  cor.coef.name = "r",
                  r.digits = 2,
                  p.digits = 2,
                  size= 5,
                  ) +
     theme_classic()  + 
     theme(axis.text.x = element_text(colour="black", size = 16)) +
     theme(axis.text.y = element_text(colour="black",size = 16)) + 
     theme(plot.title = element_text(colour="black", size=16,hjust = 0)) +
     theme(axis.title.x = element_text(colour="black", size =16)) +
     theme(axis.title.y = element_text(colour="black", size=16)) +
     theme(strip.text = element_text(size=16)) +
     theme(legend.title = element_text(color = "black", size = 12),
                  legend.text = element_text(color = "black", size = 12),
                  legend.position = "none",
                  legend.direction="vertical",
                  legend.background = element_rect(size=0.5, linetype="solid", colour ="black"))

ggsave("pop_plot.png", pop_plot, width=4, height=3)
ggsave("pop_plot.svg", pop_plot, width=4, height=3)


```
#gdp  graph 
```{r}

gdp_plot = global_compile %>% filter(!is.na(n)) %>%
  ggplot(., aes(x=gdp_2023, y=n)) +
     geom_smooth(method="lm", aes(colour=NULL, shape=NULL), size = 2.25, colour="grey60", se =FALSE)+
     geom_point(size = 4,stroke = 1,colour="black",shape=21, fill= alpha("#2b94db", 0.4)) + 
     ggtitle("GDP vs Study (n)") +
     ylab("# Studies") +
     xlab("GDP 2023 (USD)") +
     stat_cor(inherit.aes=FALSE,
                  method="pearson", 
                  data = global_compile,
                  aes(x = gdp_2023, y =n),
                  cor.coef.name = "r",
                  r.digits = 2,
                  p.digits = 2,
                  size= 5,
                  ) +
     theme_classic()  + 
     theme(axis.text.x = element_text(colour="black", size = 16)) +
     theme(axis.text.y = element_text(colour="black",size = 16)) + 
     theme(plot.title = element_text(colour="black", size=16,hjust = 0)) +
     theme(axis.title.x = element_text(colour="black", size =16)) +
     theme(axis.title.y = element_text(colour="black", size=16)) +
     theme(strip.text = element_text(size=16)) +
     theme(legend.title = element_text(color = "black", size = 12),
                  legend.text = element_text(color = "black", size = 12),
                  legend.position = "none",
                  legend.direction="vertical",
                  legend.background = element_rect(size=0.5, linetype="solid", colour ="black"))

ggsave("gdp_plot.png", gdp_plot, width=4, height=3)
ggsave("gdp_plot.svg", gdp_plot, width=4, height=3)

```

#CHEgdp  graph 
```{r}

chegdp_plot = global_compile %>% filter(!is.na(n)) %>%
  ggplot(., aes(x=che_gdp, y=n)) +
     geom_smooth(method="lm", aes(colour=NULL, shape=NULL), size = 2.25, colour="grey60", se =FALSE)+
     geom_point(size = 4,stroke = 1,colour="black",shape=21, fill= alpha("#2b94db", 0.4)) + 
     ggtitle("CHE GDP (%) vs Study (n)") +
     ylab("# Studies") +
     xlab("CHE GDP (%)") +
     stat_cor(inherit.aes=FALSE,
                  method="pearson", 
                  data = global_compile,
                  aes(x = che_gdp, y =n),
                  cor.coef.name = "r",
                  r.digits = 2,
                  p.digits = 2,
                  size= 5,
                  ) +
     theme_classic()  + 
     theme(axis.text.x = element_text(colour="black", size = 16)) +
     theme(axis.text.y = element_text(colour="black",size = 16)) + 
     theme(plot.title = element_text(colour="black", size=16,hjust = 0)) +
     theme(axis.title.x = element_text(colour="black", size =16)) +
     theme(axis.title.y = element_text(colour="black", size=16)) +
     theme(strip.text = element_text(size=16)) +
     theme(legend.title = element_text(color = "black", size = 12),
                  legend.text = element_text(color = "black", size = 12),
                  legend.position = "none",
                  legend.direction="vertical",
                  legend.background = element_rect(size=0.5, linetype="solid", colour ="black"))

ggsave("chegdp_plot.png", chegdp_plot, width=4, height=3)
ggsave("chegdp_plot.svg", chegdp_plot, width=4, height=3)

```


#**Country data with surg_onc vs other 

323 codes from your data successfully matched countries in the map
22 codes from your data failed to match with a country code in the map
      failedCodes failedCountries                       
 [1,] ""          "Bonaire, Sint Eustatius and Saba"    
 [2,] NA          "Czechia"                             
 [3,] NA          "Côte D'Ivoire"                       
 [4,] NA          "Former Serbia and Montenegro"        
 [5,] NA          "Former Yugoslavia"                   
 [6,] "GIB"       "Gibraltar"                           
 [7,] "GLP"       "Guadeloupe"                          
 [8,] NA          "Libyan Arab Jamahiriya"              
 [9,] "MTQ"       "Martinique"                          
[10,] "MYT"       "Mayotte"                             
[11,] ""          "Netherlands Antilles"                
[12,] NA          "North Macedonia"                     
[13,] NA          "Palestinian Territories, Occupied"   
[14,] NA          "Réunion"                             
[15,] "UMI"       "United States Minor Outlying Islands"
[16,] NA          NA                                    
[17,] NA          "Czechia"                             
[18,] NA          "Former Serbia and Montenegro"        
[19,] NA          "Libyan Arab Jamahiriya"              
[20,] "MTQ"       "Martinique"                          
[21,] NA          "North Macedonia"                     
[22,] NA          NA                                    
36 codes from the map weren't represented in your data

```{r}
ensemble_countries_all = merge(ensemble_FINAL_surgonc,countries, by="nct_id")
ensemble_countries_all$name = str_replace_all(ensemble_countries_all$name, "Palestinian Territories, Occupied", "Palestinian Territory, Occupied")
ensemble_countries_all = ensemble_countries_all[!is.na(ensemble_countries_all$name),]
ensemble_countries_all = ensemble_countries_all %>% group_by(class) %>% distinct(nct_id, name) %>% count(name)

global_compile_all = merge(world_bank_compile, WHO_healthindicators_filtered_final, by.x = "Country_Code", by.y = "DIM_GEO_CODE",all.x = TRUE, all.y = TRUE)

global_compile_all = merge(global_compile_all, Health_expenditures_filtered, by.x = "Country_Code", by.y = "code",all.x = TRUE, all.y = TRUE)

global_compile_all = merge(global_compile_all, health_cluster, by.x = "Country_Code", by.y = "Country_Code", all.x = TRUE, all.y = TRUE)




ensemble_countries_all$name = str_replace_all(ensemble_countries_all$name, c("Czechia" = "Czech Republic"))
ensemble_countries_all$name = str_replace_all(ensemble_countries_all$name, c("Former Serbia and Montenegro" = "Serbia"))
ensemble_countries_all$name = str_replace_all(ensemble_countries_all$name, c("Libyan Arab Jamahiriya" = "Libya"))
ensemble_countries_all$name = str_replace_all(ensemble_countries_all$name, c("Martinique" = "France"))
ensemble_countries_all$name = str_replace_all(ensemble_countries_all$name, c("North Macedonia" = "Macedonia"))

ensemble_countries_all$name = str_replace_all(ensemble_countries_all$name, c("Côte D'Ivoire" = "Cote d'Ivoire"))
ensemble_countries_all$name = str_replace_all(ensemble_countries_all$name, c("Gibraltar" = "Gibraltar"))
#ensemble_countries_all$name = str_replace_all(ensemble_countries_all$name, c("Territorial Collectivity of Mayotte" = "Mayotte"))
ensemble_countries_all$name = str_replace_all(ensemble_countries_all$name, c("Netherlands Antilles" = "Netherlands"))
ensemble_countries_all$name = str_replace_all(ensemble_countries_all$name, c("Réunion" = "France"))
ensemble_countries_all$name = str_replace_all(ensemble_countries_all$name, c("Guadeloupe" = "Department of Guadeloupe"))

#Check the total n studies 
sum(ensemble_countries_all$n)
#Make sure there are no duplicates by class
ensemble_countries_all <- ensemble_countries_all %>%
  group_by(class, name) %>%
  summarise(n = sum(n), .groups = "drop")
#Check the total n studies (make sure they are consistent)
sum(ensemble_countries_all$n)

names_match = dplyr::select(ensemble_countries_all, "name") #Name from the ensemble set

country_list = as.data.frame(countrySynonyms)  #Country list synonyms
country_list$ISO3 <- toupper(country_list$ISO3) #Make upper case

country_list_long <- country_list %>% 
  tidyr::pivot_longer(cols = starts_with("name"), names_to = "name_type", values_to = "name") %>%
  select(-name_type) #Perfirm this to map country names to and ISO3

countryname_merged <- names_match %>%
  left_join(country_list_long, by = c("name" = "name"))
countryname_merged <- countryname_merged %>%
  distinct(name, .keep_all = TRUE)
countryname_merged = countryname_merged %>% filter(!is.na(name)) 


ensemble_countries_all = merge(countryname_merged, ensemble_countries_all, by.x ="name", by.y="name", all.x = TRUE, all.y = TRUE)
#Check the total n studies 
sum(ensemble_countries_all$n)
######


global_compile_all = merge(ensemble_countries_all, global_compile_all, by.x = "ISO3", by.y = "Country_Code", all.x = TRUE, all.y = TRUE)

global_compile_all = global_compile_all %>% filter(!is.na(name)) 

global_compile_all = dplyr::select(global_compile_all, c("ISO3", "n", "name","class", "Region", "IncomeGroup", "pop_2000", "gdp_2023","pop_2020", "pop_2023", "Life_expectancy_at_birth_(years)", "Total_net_official_development_assistance_to_medical_research_and_basic_health_sectors_per_capita_(US$),_by_recipient_country_", "Density_of_medical_doctors_(per_10_000_population)_", "Population_with_household_expenditures_on_health_>_25%_of_total_household_expenditure_or_income_(%)" ,"year",  "che_gdp", "hc71_usd", "SHI_as_%_THE","Government_financing_as_%_of_THE","OOP_as_%_of_THE","VHI_as_%_of_THE","Predominance_defined_by_clustering", "Predominance_as_\"largest_HFS\""   ))

economy = dplyr::select(country_mapping_world, c("name.y", "economy"))

global_compile_all = merge(global_compile_all, economy, by.x = "name", by.y = "name.y", all.x = TRUE, all.y = TRUE)

sum(global_compile_all$n) #It is different because we lost NA data. (But it is less which means there was not duplicates created)

```

#Make compiled long
```{r}
global_compile_all


df_expanded <- global_compile_all %>%
  tidyr::uncount(weights = n)
```

#Tableone ECONOMIC AND GLOBAL HEALTH DATA TABLES
```{r}
library(tableone)
library(flextable)
library(gtsummary)

## define custom test #############
fisher.test.simulate.p.values <- function(data, variable, by, ...) {
  result <- list()
  test_results <- stats::fisher.test(data[[variable]], data[[by]], simulate.p.value = TRUE)
  result$p <- test_results$p.value
  result$test <- test_results$method
  result
}

 tb1 = tbl_summary(global_compile_all,
             by = "class", 
             percent = "column", 
             include=c("n", "pop_2000", "gdp_2023", "pop_2023", "Life_expectancy_at_birth_(years)", "Total_net_official_development_assistance_to_medical_research_and_basic_health_sectors_per_capita_(US$),_by_recipient_country_", "che_gdp",  "SHI_as_%_THE","Government_financing_as_%_of_THE","OOP_as_%_of_THE","VHI_as_%_of_THE","Predominance_defined_by_clustering", "Predominance_as_\"largest_HFS\"")) %>%  add_p() %>% add_n(col_label="**n/N (Missing %)**", statistic= "{N_miss} / {N_obs} ({p_miss}%)", add_footnote = "Percent") 

 tb1_flex = as_flex_table(
  tb1,
  include = everything(),
  return_calls = FALSE,
  strip_md_bold = TRUE
)

save_as_docx("surg_onc_table1" = tb1_flex,path = "/Users/danielskubleny/Documents/R projects/tensorflow_probs3/surg_onc_table1.docx")


####LONG



 tb2 = tbl_summary(df_expanded,
             by = "class", 
             percent = "column", 
             include=c("Region", "IncomeGroup", "Predominance_defined_by_clustering", "Predominance_as_\"largest_HFS\"")) %>%  add_p() %>% add_n(col_label="**n/N (Missing %)**", statistic= "{N_miss} / {N_obs} ({p_miss}%)", add_footnote = "Percent") 

 tb2_flex = as_flex_table(
  tb2,
  include = everything(),
  return_calls = FALSE,
  strip_md_bold = TRUE
)
 
 save_as_docx("surg_onc_table2" = tb2_flex,path = "/Users/danielskubleny/Documents/R projects/tensorflow_probs3/surg_onc_table2.docx")
 
 


```

#Compile common dataset for other data with no duplicate nct_id
```{r}
eligibilities = dplyr::select(eligibilities, c("nct_id", "gender", "minimum_age", "maximum_age", "healthy_volunteers"))

studies = dplyr::select(studies, c("nct_id", "study_first_submitted_date", "results_first_submitted_date", "last_update_submitted_date", "study_first_posted_date", "last_update_submitted_qc_date", "study_type", "overall_status", "phase", "enrollment", "enrollment_type", "source", "number_of_arms", "number_of_groups", "why_stopped", "has_expanded_access", "is_fda_regulated_device", "is_fda_regulated_drug", "is_unapproved_device", "is_us_export", "biospec_retention", "plan_to_share_ipd", "source_class"))

calculated_values = dplyr::select(calculated_values, c("nct_id", "number_of_facilities", "registered_in_calendar_year", "actual_duration", "were_results_reported", "months_to_report_results", "has_us_facility", "has_single_facility", "number_of_primary_outcomes_to_measure", "number_of_secondary_outcomes_to_measure"))

designs = dplyr::select(designs, c("nct_id", "allocation", "intervention_model", "observational_model", "primary_purpose", "time_perspective", "masking", "subject_masked", "caregiver_masked", "investigator_masked", "outcomes_assessor_masked"))

combo_data = merge(studies, designs, by="nct_id", all.x=TRUE, all.y=TRUE)
combo_data = merge(combo_data, eligibilities, by="nct_id", all.x=TRUE, all.y=TRUE)
combo_data = merge(combo_data, calculated_values, by="nct_id", all.x=TRUE, all.y=TRUE)

surg_onc_data = dplyr::select(ensemble_FINAL_surgonc, c("nct_id", "class", "mean"))

combo_data = merge(surg_onc_data, combo_data, by="nct_id", all.x=TRUE)

#FIX "number_of_primary_outcomes_to_measure", "number_of_secondary_outcomes_to_measure", "number_of_arms", "number_of_groups", "actua duration#####

combo_data$number_of_facilities = as.numeric(combo_data$number_of_facilities)


plot(as.factor(combo_data$class),as.numeric(combo_data$months_to_report_results))
table(combo_data$class, combo_data$allocation)

plot(as.factor(combo_data$class),as.numeric(combo_data$number_of_facilities))

```
#Network analysis 
```{r}
sponsors_network = merge(ensemble_FINAL_surgonc, sponsors, by="nct_id")
sponsors_network = sponsors_network %>% filter(class=="yes")
```
There are issues with some of the names. For example Novartis and Novartis Pharmaceuticals is the same institution. I need to identify these instances and adjust names accordingly. 
```{r}
library(stringdist)
institutions = sponsors_network$name
# Compute a string distance matrix
dist_matrix <- stringdistmatrix(institutions, institutions, method = "jw")  # Jaro-Winkler
```
```{r}
# Convert to a data frame for easier viewing
similarities <- which(dist_matrix < 0.15 & dist_matrix > 0, arr.ind = TRUE)
results <- unique(data.frame(
  name1 = institutions[similarities[,1]],
  name2 = institutions[similarities[,2]],
  distance = dist_matrix[similarities]
))

```
#Replace strings
I assessed the closely related/misspelled strings and exchange those where there are duplicates. I maintained distinctions to sepraate campus'/sub-institutions where applicable. 
```{r}
sponsors_network$name = str_replace_all(sponsors_network$name, "National Institute of Diabetes and Digestive and Kidney Diseases (NIDDK)", "National Institute for Diabetes and Digestive and Kidney Diseases (NIDDK)")
sponsors_network$name = str_replace_all(sponsors_network$name, "Guangzhou University of Traditional Chinese Medicine", "Guangzhou University of Chinese Medicine")
sponsors_network$name = str_replace_all(sponsors_network$name, "Memorial Sloan Kettering Cancer Center, New York, USA", "Memorial Sloan Kettering Cancer Center")
sponsors_network$name = str_replace_all(sponsors_network$name, "Novartis Pharmaceuticals", "Novartis")
sponsors_network$name = str_replace_all(sponsors_network$name, "Beijing Tsinghua Chang Gung Hospital", "Beijing Tsinghua Changgeng Hospital")
sponsors_network$name = str_replace_all(sponsors_network$name, "Universtiy of Mississippi Medical Center", "University of Mississippi Medical Center")
sponsors_network$name = str_replace_all(sponsors_network$name, "Istanbul Training and Researh Hospital", "Istanbul Training and Research Hospital")
sponsors_network$name = str_replace_all(sponsors_network$name, "Tannaz Armaghnay", "Tannaz Armaghany")
sponsors_network$name = str_replace_all(sponsors_network$name, "NeoImmune Tech", "NeoImmuneTech")
sponsors_network$name = str_replace_all(sponsors_network$name, "Carlos Cerdán Santacruz", "Carlos Cerdán-Santacruz")
sponsors_network$name = str_replace_all(sponsors_network$name, "University Hospital Padua", "University Hospital Padova")
sponsors_network$name = str_replace_all(sponsors_network$name, "European Organisation for Research and Treatment of Cancer", "European Organisation for Research and Treatment of Cancer - EORTC")
sponsors_network$name = str_replace_all(sponsors_network$name, "Ismail Gögenür", "Ismail Gögenur")
sponsors_network$name = str_replace_all(sponsors_network$name, "Translational Breast Cancer Research Consortium", "Translational Breast Cancer Research Consortium (TBCRC)")
sponsors_network$name = str_replace_all(sponsors_network$name, "Austrian Breast & Colorectal Cancer Study Group", "Austrian Breast & Colorectal Cancer Study Group (ABCSG)")
sponsors_network$name = str_replace_all(sponsors_network$name, "Università Cattolica del Sacro Cuore-IRCCS, Rome, Italy", "Università Cattolica del Sacro Cuore-IRCCS Rome")
sponsors_network$name = str_replace_all(sponsors_network$name, "YuceBio Technology Co., Ltd.", "Yuce Biotechnology Co., Ltd.")
sponsors_network$name = str_replace_all(sponsors_network$name, "Tianjin Medical University Cancer Institute & Hospital", "Tianjin Medical University Cancer Institute and Hospital")
sponsors_network$name = str_replace_all(sponsors_network$name, "Chinese Cooperative Group of Liver Cancer", "Chinese Cooperative Group of Liver Cancer (CCGLC)")
sponsors_network$name = str_replace_all(sponsors_network$name, "Tokyo Medical and Dental University, Japan", "Tokyo Medical and Dental University")
sponsors_network$name = str_replace_all(sponsors_network$name, "Vall d'Hebron Institute of Oncology", "Vall d'Hebron Institute of Oncology (VHIO)")
sponsors_network$name = str_replace_all(sponsors_network$name, "Clinica Universidad de Navarra, Pamplona, Spain", "Clinica Universidad de Navarra Pamplona")
sponsors_network$name = str_replace_all(sponsors_network$name, "Spanish Breast Cancer Research Group", "Spanish Breast Cancer Research Group (GEICAM)")
sponsors_network$name = str_replace_all(sponsors_network$name, "SOLTI Breast Cancer Research Group", "Spanish Breast Cancer Research Group (SOLTI)")
sponsors_network$name = str_replace_all(sponsors_network$name, "Umberto I Mauriziano Hospital, Turin, Italy", "Umberto I Mauriziano Hospital Turin")
sponsors_network$name = str_replace_all(sponsors_network$name, "Kantonsspital Graubuenden", "Kantonsspital Graubünden")
sponsors_network$name = str_replace_all(sponsors_network$name, "Dutch Colorectal Cancer Group", "Dutch Colorectal Cancer Group (DCCG)")
sponsors_network$name = str_replace_all(sponsors_network$name, "Dutch Pancreatic Cancer Group", "Dutch Pancreatic Cancer Group (DPCG)")
sponsors_network$name = str_replace_all(sponsors_network$name, "Sunnybrook Health Sciences Centre, Canada", "Sunnybrook Health Sciences Centre")
sponsors_network$name = str_replace_all(sponsors_network$name, "IRCCS San Raffaele Hospital, Milan, Italy", "IRCCS San Raffaele Hospital Milan")
sponsors_network$name = str_replace_all(sponsors_network$name, "West China School of Medicine and West China Hospital, Sichuan University", "West China School of Medicine/West China Hospital of Sichuan University")
sponsors_network$name = str_replace_all(sponsors_network$name, "Oncomédics", "Oncomedics")
sponsors_network$name = str_replace_all(sponsors_network$name, "Jianjun Yang", "Jianjun Yang,MD")
sponsors_network$name = str_replace_all(sponsors_network$name, "Italian Research Group for Gastric Cancer", "Italian Research Group For Gastric Cancer - IRGGC")
sponsors_network$name = str_replace_all(sponsors_network$name, "Herbert Hurwitz", "Herbert Hurwitz, MD")
sponsors_network$name = str_replace_all(sponsors_network$name, "Bryan Schneider", "Bryan Schneider, MD")
sponsors_network$name = str_replace_all(sponsors_network$name, "Joseph J. Cullen", "Joseph J. Cullen, MD, FACS")
sponsors_network$name = str_replace_all(sponsors_network$name, "Steven J Isakoff, MD, PhD", "Steven Isakoff, MD, PhD")
sponsors_network$name = str_replace_all(sponsors_network$name, "Xinrui Zhu", "Xinrui Zhu，MD")
sponsors_network$name = str_replace_all(sponsors_network$name, "Howard S Hochster", "Howard S. Hochster, MD")
sponsors_network$name = str_replace_all(sponsors_network$name, "Korean Cancer Study Group", "Korean Cancer Study Group (KCSG)")
sponsors_network$name = str_replace_all(sponsors_network$name, "Institut d'Investigacions Biomèdiques August Pi i Sunyer", "Institut d'Investigacions Biomèdiques August Pi i Sunyer (IDIBAPS), Spain")


```

#Interactive Plot NEW

Here I will have the code that is pasted into and app.R script file and placed to my final directory. 

```{r}
#Need to have this placed into my shiny app folder. 
saveRDS(sponsors_network, "sponsors_network.rds")

```
```{r}
library(shiny)
library(shinydashboard)
library(networkD3)
library(igraph)
library(tidygraph)
library(dplyr)
library(htmlwidgets)
library(htmltools)

sponsors_network <- readRDS("data/sponsors_network.rds")


# Step 1: Identify lead → collaborator edges
edge_directed <- sponsors_network %>%
  dplyr::filter(lead_or_collaborator == "lead") %>%
  dplyr::rename(lead = name) %>%
  dplyr::select(nct_id, lead) %>%
  dplyr::inner_join(
    sponsors_network %>% dplyr::filter(lead_or_collaborator == "collaborator"),
    by = "nct_id"
  ) %>%
  dplyr::select(from = lead, to = name)

# Step 2: Optionally count how often each lead → collaborator pair appears
edge_weighted <- edge_directed %>%
  count(from, to, name = "weight")

# Keep only people who have led at least 60 times
important_people <- sponsors_network %>%
  filter(lead_or_collaborator == "lead") %>%
  count(name) %>%
  filter(n >= 60) %>%  # change threshold as needed
  pull(name)

# Create directed igraph object
g_directed <- graph_from_data_frame(edge_weighted, directed = TRUE)

# Filter graph to only include those
g_directed <- induced_subgraph(g_directed, vids = V(g_directed)$name %in% important_people)

g_directed = simplify(g_directed)

####OPTION TO MAKE UNDIRECTED FOR CLUSTER
# Convert to tbl_graph (still directed)
tg_directed <- as_tbl_graph(g_directed)

# Create the tidygraph object with Louvain community info
tg_directed <- tg_directed %>%
  mutate(
    degree = centrality_degree(mode = "all"),
    degree_in = centrality_degree(mode = "in"),
    degree_out = centrality_degree(mode = "out"),
    authority = centrality_authority(weights = weight, scale = FALSE),
    eigen = centrality_eigen(weights = weight, directed=TRUE,scale = FALSE),
    pargerank = centrality_pagerank(weights = weight, directed=TRUE),
    betweenness = centrality_betweenness())

tg_directed <- tg_directed %>%
  activate(nodes) %>%
  filter(degree > 5)  # Filter by degree and community

# Convert to undirected and calculate community
g_undirected <- as_undirected(tg_directed, mode = "collapse")
louvain_clusters <- cluster_louvain(g_undirected)

tg_directed <- tg_directed %>%
  mutate(community = as.factor(membership(louvain_clusters)[name]))

# Step 1: Identify lead → collaborator edges
edge_directed <- sponsors_network %>%
  dplyr::filter(lead_or_collaborator == "lead") %>%
  dplyr::rename(lead = name) %>%
  dplyr::select(nct_id, lead) %>%
  dplyr::inner_join(
    sponsors_network %>% dplyr::filter(lead_or_collaborator == "collaborator"),
    by = "nct_id"
  ) %>%
  dplyr::select(from = lead, to = name)

# Step 2: Optionally count how often each lead → collaborator pair appears
edge_weighted <- edge_directed %>%
  count(from, to, name = "weight")



# Create directed igraph object
g_directed <- graph_from_data_frame(edge_weighted, directed = TRUE)

g_directed = simplify(g_directed)

####OPTION TO MAKE UNDIRECTED FOR CLUSTER
# Convert to tbl_graph (still directed)
tg_directed <- as_tbl_graph(g_directed)

# Create the tidygraph object with Louvain community info
tg_directed <- tg_directed %>%
  mutate(
    degree = centrality_degree(mode = "all"),
    degree_in = centrality_degree(mode = "in"),
    degree_out = centrality_degree(mode = "out"),
    authority = centrality_authority(weights = weight, scale = FALSE),
    eigen = centrality_eigen(weights = weight, directed=TRUE,scale = FALSE),
    pargerank = centrality_pagerank(weights = weight, directed=TRUE),
    betweenness = centrality_betweenness())

# Convert to undirected and calculate community
g_undirected <- as_undirected(tg_directed, mode = "collapse")
set.seed(99)
louvain_clusters_large <- cluster_louvain(g_undirected)

tg_directed <- tg_directed %>%
  mutate(community = as.factor(membership(louvain_clusters_large)[name]))


# Step 1: Identify lead → collaborator edges
edge_directed <- sponsors_network %>%
  dplyr::filter(lead_or_collaborator == "lead") %>%
  dplyr::rename(lead = name) %>%
  dplyr::select(nct_id, lead) %>%
  dplyr::inner_join(
    sponsors_network %>% dplyr::filter(lead_or_collaborator == "collaborator"),
    by = "nct_id"
  ) %>%
  dplyr::select(from = lead, to = name)

# Step 2: Optionally count how often each lead → collaborator pair appears
edge_weighted <- edge_directed %>%
  count(from, to, name = "weight")

# Create directed igraph object
g_directed <- graph_from_data_frame(edge_weighted, directed = TRUE)

g_directed = simplify(g_directed)

####OPTION TO MAKE UNDIRECTED FOR CLUSTER
# Convert to tbl_graph (still directed)
tg_directed <- as_tbl_graph(g_directed)

# Create the tidygraph object with Louvain community info
tg_directed <- tg_directed %>%
  mutate(
    degree = centrality_degree(mode = "all"),
    degree_in = centrality_degree(mode = "in"),
    degree_out = centrality_degree(mode = "out"),
    authority = centrality_authority(weights = weight, scale = FALSE),
    eigen = centrality_eigen(weights = weight, directed=TRUE,scale = FALSE),
    pargerank = centrality_pagerank(weights = weight, directed=TRUE),
    betweenness = centrality_betweenness())

tg_directed <- tg_directed %>%
  mutate(community = as.factor(membership(louvain_clusters_large)[name]))

# Step 3: Join colors back into the tidygraph object
tg_directed <- tg_directed %>%
  activate(nodes) %>%
  left_join(community_colors, by = "community")

# Step 1: Compute connected components (works for directed or undirected)
tg_filtered <- tg_directed %>%
  mutate(component = group_components())  # assigns a component ID to each node

# Step 2: Identify the largest component by number of nodes
largest_component <- tg_filtered %>%
  as_tibble() %>%
  count(component, sort = TRUE) %>%
  slice(1) %>%
  pull(component)

# Step 3: Filter the graph to only keep nodes in the largest component
tg_main_component <- tg_filtered %>%
  activate(nodes) %>%
  filter(component == largest_component)

important_graph_names = tg_directed_important %>% 
  as_tibble() %>% 
  pull(name)

tg_main_component_check <- tg_filtered %>%
  activate(nodes) %>%
  filter(name %in% important_graph_names)
tg_main_component_check = tg_main_component_check %>% as_tibble()
tg_main_component_check = as.data.frame(tg_main_component_check)


#Filter by institutions with at least 60 Leads
tg_directed_subgraph <- tg_filtered %>%
  activate(nodes) %>%
  filter(name %in% important_people)

tg_directed_subgraph <- tg_directed_subgraph %>%
  mutate(
    degree = centrality_degree(mode = "all"),
    degree_in = centrality_degree(mode = "in"),
    degree_out = centrality_degree(mode = "out"),
    authority = centrality_authority(weights = weight, scale = FALSE),
    eigen = centrality_eigen(weights = weight, directed=TRUE,scale = FALSE),
    pargerank = centrality_pagerank(weights = weight, directed=TRUE),
    betweenness = centrality_betweenness())

#Filter by institutions with degree > 5 due to disconnected graph and too many communities for communication 
tg_directed_subgraph <- tg_directed_subgraph %>%
  activate(nodes) %>%
  filter(degree > 5)  # Filter by degree and community

# Extract nodes and links
nodes <- tg_directed_subgraph %>%
  activate(nodes) %>%
  as_tibble() %>%
  mutate(id = row_number() - 1)  # forceNetwork uses 0-based indexing
nodes = as.data.frame(nodes)
nodes$degree_exp = nodes$degree^1.4

edges <- tg_directed_subgraph %>%
  activate(edges) %>%
  as_tibble() %>%
  mutate(
    source = as.integer(from) - 1,  # 0-based indexing
    target = as.integer(to) - 1
  )
edges = as.data.frame(edges)

color_scale <- JS("d3.scaleOrdinal()
                    .domain([1, 11, 16, 25, 28,  5,  6,  7 ])
                    .range(['#33A02C', '#FDBF6F', '#CAB2D6','#2b94db', '#FF7F00', '#B2DF8A', '#6A3D9A', '#A6CEE3']);")

#EXPORT TO SHINY
# ==== UI ====

ui <- dashboardPage(
  dashboardHeader(title = "Network Surgical Oncology"),
  dashboardSidebar(
    sidebarMenu(
      menuItem("Interactive Network", tabName = "network", icon = shiny::icon("project-diagram")),
      shiny::textInput("search", "Search Node:", "")
    )
  ),
  dashboardBody(
    tabItems(
      tabItem(tabName = "network",
              box(title = "Force Network Graph", width = 12, status = "primary",
                  forceNetworkOutput("net", height = "700px")
              )
      )
    )
  )
)

# ==== Server ====

server <- function(input, output, session) {
  output$net <- renderForceNetwork({
    fn <- forceNetwork(
      Links = edges,
      Nodes = nodes,
      Source = "source",
      Target = "target",
      NodeID = "name",
      Group = "community",   # color by community
      Value = "weight",      # edge width based on weight
      opacity = 0.9,
      zoom = TRUE,
      linkDistance = JS("function(d) { return 50 / d.value; }"),  # shorter distance for heavier weight
      fontSize = 12,
      linkColour = "#DCDCDC",
      charge = -500,
      Nodesize = "degree_exp",       # or another numeric column
      colourScale = color_scale
    )
    
    # Add visible labels and search behavior
    onRender(fn, '
function(el,x) { 
    var link = d3.selectAll(".link")
    var node = d3.selectAll(".node")

    var options = { opacity: 1,
                    clickTextSize: 10,
                    opacityNoHover: 0.1,
                    radiusCalculation: "Math.sqrt(d.nodesize)+6"
                  }

    var unfocusDivisor = 4;

    var links = HTMLWidgets.dataframeToD3(x.links);
    var linkedByIndex = {};

    links.forEach(function(d) {
      linkedByIndex[d.source + "," + d.target] = 1;
      linkedByIndex[d.target + "," + d.source] = 1;
    });

    function neighboring(a, b) {
      return linkedByIndex[a.index + "," + b.index];
    }

    function nodeSize(d) {
            if(options.nodesize){
                    return eval(options.radiusCalculation);
            }else{
                    return 6}
    }

    function mouseover(d) {
      var unfocusDivisor = 4;

      link.transition().duration(200)
        .style("opacity", function(l) { return d != l.source && d != l.target ? +options.opacity / unfocusDivisor : +options.opacity });

      node.transition().duration(200)
        .style("opacity", function(o) { return d.index == o.index || neighboring(d, o) ? +options.opacity : +options.opacity / unfocusDivisor; });

      d3.select(this).select("circle").transition()
        .duration(750)
        .attr("r", function(d){return nodeSize(d)+5;});

      node.select("text").transition()
        .duration(750)
        .attr("x", 13)
        .style("stroke-width", ".5px")
        .style("font", 24 + "px ")
        .style("opacity", function(o) { return d.index == o.index || neighboring(d, o) ? 1 : 0; });
    }

    function mouseout() {
      node.style("opacity", +options.opacity);
      link.style("opacity", +options.opacity);

      d3.select(this).select("circle").transition()
        .duration(750)
        .attr("r", function(d){return nodeSize(d);});
      node.select("text").transition()
        .duration(1250)
        .attr("x", 0)
        .style("font", options.fontSize + "px ")
        .style("opacity", 0);
    }

    d3.selectAll(".node").on("mouseover", mouseover).on("mouseout", mouseout);
    // Add search highlighting
    var input = document.getElementById("search");
    input.addEventListener("input", function() {
      var searchTerm = this.value.toLowerCase();
      d3.select(el).selectAll(".node circle")
        .style("stroke", function(d) {
          return d.name.toLowerCase().includes(searchTerm) ? "red" : null;
        })
        .style("stroke-width", function(d) {
          return d.name.toLowerCase().includes(searchTerm) ? 3 : null;
        });
    });
}
')
  })
}

# ==== Run App ====
shinyApp(ui = ui, server = server)
```
#Sort out the working directories and deploy app
```{r}
original_wd = getwd()

setwd("/Users/danielskubleny/Documents/app/surg_onc_network_plot/network-surgical-oncology")  # Or wherever your app.R is located

rsconnect::deployApp("/Users/danielskubleny/Documents/app/surg_onc_network_plot/network-surgical-oncology")

```
#Deploy app

shiny::runApp("/Users/danielskubleny/Documents/app/surg_onc_network_plot/network-surgical-oncology")


library(rsconnect)
rsconnect::deployApp('/Users/danielskubleny/Documents/app/surg_onc_network_plot/network-surgical-oncology')  # e.g., "C:/Users/you/Documents/myApp"

###https://3j03cb-daniel0skubleny.shinyapps.io/network-surgical-oncology/####



#SOCIAL NETWORK LARGE for degree in and degree out 
```{r}
# Step 1: Identify lead → collaborator edges
edge_directed <- sponsors_network %>%
  dplyr::filter(lead_or_collaborator == "lead") %>%
  dplyr::rename(lead = name) %>%
  dplyr::select(nct_id, lead) %>%
  dplyr::inner_join(
    sponsors_network %>% dplyr::filter(lead_or_collaborator == "collaborator"),
    by = "nct_id"
  ) %>%
  dplyr::select(from = lead, to = name)

# Step 2: Optionally count how often each lead → collaborator pair appears
edge_weighted <- edge_directed %>%
  count(from, to, name = "weight")



# Create directed igraph object
g_directed <- graph_from_data_frame(edge_weighted, directed = TRUE)

g_directed = simplify(g_directed)

####OPTION TO MAKE UNDIRECTED FOR CLUSTER
# Convert to tbl_graph (still directed)
tg_directed <- as_tbl_graph(g_directed)

# Create the tidygraph object with Louvain community info
tg_directed <- tg_directed %>%
  mutate(
    degree = centrality_degree(mode = "all"),
    degree_in = centrality_degree(mode = "in"),
    degree_out = centrality_degree(mode = "out"),
    authority = centrality_authority(weights = weight, scale = FALSE),
    eigen = centrality_eigen(weights = weight, directed=TRUE,scale = FALSE),
    pargerank = centrality_pagerank(weights = weight, directed=TRUE),
    betweenness = centrality_betweenness())

# Convert to undirected and calculate community
g_undirected <- as_undirected(tg_directed, mode = "collapse")
set.seed(99)
louvain_clusters_large <- cluster_louvain(g_undirected)

tg_directed <- tg_directed %>%
  mutate(community = as.factor(membership(louvain_clusters_large)[name]))

# Step 1: Extract and shuffle unique communities
set.seed(123)  # for reproducibility
unique_communities <- tg_directed %>%
  activate(nodes) %>%
  as_tibble() %>%
  pull(community) %>%
  unique() %>%
  sample()  # randomize order

# Step 2: Assign random colors (from viridis) to each community
community_colors <- data.frame(
  community = unique_communities,
  color = viridis(length(unique_communities))
)

# Your known subset of 8 communities you want to assign fixed colors to
target_communities <- c("7",  "25", "5",  "1",  "11", "28", "16", "6")  # Replace with your actual communities

# Generate viridis(8) palette
fixed_colors <- brewer.pal(10,"Paired")
fixed_colors <- fixed_colors[-c(5:6)]

# Step 1: Create a named vector for overrides
override_colors <- setNames(fixed_colors, target_communities)

# Step 2: Update `community_colors` with the overrides
community_colors <- community_colors %>%
  mutate(community = as.character(community)) %>%
  mutate(color = ifelse(community %in% names(override_colors),
                        override_colors[as.character(community)],
                        color))

target_second_round = "25"
override_cols = "#2b94db"

override_second <- setNames(override_cols, target_second_round)

# Step 2b: Update `community_colors` with the overrides
community_colors <- community_colors %>%
  mutate(community = as.character(community)) %>%
  mutate(color = ifelse(community %in% names(override_second),
                        override_second[as.character(community)],
                        color))

        
# Step 3: Join colors back into the tidygraph object
tg_directed <- tg_directed %>%
  activate(nodes) %>%
  left_join(community_colors, by = "community")

set.seed(99)
mds_large = ggraph(tg_directed, layout = "mds") +  # Fruchterman-Reingold layout
  geom_edge_link0(aes(width = weight), alpha = 0.3) +
  geom_node_point(aes(size = degree, fill = I(color)), shape=21, alpha=0.1) + 
  scale_edge_width(name = "Weight",range = c(0.1, 1)) +
  scale_size_continuous(name = "Degree", range = c(3, 7)) + 
  labs(title = "Complete Network", y="Dim 2", x = "Dim 1") +
  theme_classic() + 
     theme(axis.text.x = element_text(colour="black", size = 18)) +
     theme(axis.text.y = element_text(colour="black",size = 18)) + 
     theme(plot.title = element_text(colour="black", size=18,hjust = 0)) +
     theme(axis.title.x = element_text(colour="black", size =18)) +
     theme(axis.title.y = element_text(colour="black", size=18)) + 
    theme(legend.position = "none") 

ggsave("mds_large.svg", mds_large, width=4, height=4)


network_data = tg_directed %>% as.data.frame()
network_data = network_data %>% select(name,degree,degree_in,degree_out,eigen,community)
write.csv(network_data, "network_results.csv")
```
#Social network semi large with subset graph nodes highlighted. 
```{r}
# Step 1: Identify lead → collaborator edges
edge_directed <- sponsors_network %>%
  dplyr::filter(lead_or_collaborator == "lead") %>%
  dplyr::rename(lead = name) %>%
  dplyr::select(nct_id, lead) %>%
  dplyr::inner_join(
    sponsors_network %>% dplyr::filter(lead_or_collaborator == "collaborator"),
    by = "nct_id"
  ) %>%
  dplyr::select(from = lead, to = name)

# Step 2: Optionally count how often each lead → collaborator pair appears
edge_weighted <- edge_directed %>%
  count(from, to, name = "weight")



# Create directed igraph object
g_directed <- graph_from_data_frame(edge_weighted, directed = TRUE)

g_directed = simplify(g_directed)

####OPTION TO MAKE UNDIRECTED FOR CLUSTER
# Convert to tbl_graph (still directed)
tg_directed <- as_tbl_graph(g_directed)

# Create the tidygraph object with Louvain community info
tg_directed <- tg_directed %>%
  mutate(
    degree = centrality_degree(mode = "all"),
    degree_in = centrality_degree(mode = "in"),
    degree_out = centrality_degree(mode = "out"),
    authority = centrality_authority(weights = weight, scale = FALSE),
    eigen = centrality_eigen(weights = weight, directed=TRUE,scale = FALSE),
    pargerank = centrality_pagerank(weights = weight, directed=TRUE),
    betweenness = centrality_betweenness())

tg_directed <- tg_directed %>%
  mutate(community = as.factor(membership(louvain_clusters_large)[name]))

# Step 3: Join colors back into the tidygraph object
tg_directed <- tg_directed %>%
  activate(nodes) %>%
  left_join(community_colors, by = "community")

# Step 1: Compute connected components (works for directed or undirected)
tg_filtered <- tg_directed %>%
  mutate(component = group_components())  # assigns a component ID to each node

# Step 2: Identify the largest component by number of nodes
largest_component <- tg_filtered %>%
  as_tibble() %>%
  count(component, sort = TRUE) %>%
  slice(1) %>%
  pull(component)

# Step 3: Filter the graph to only keep nodes in the largest component
tg_main_component <- tg_filtered %>%
  activate(nodes) %>%
  filter(component == largest_component)

important_graph_names = tg_directed_important %>% 
  as_tibble() %>% 
  pull(name)

tg_main_component_check <- tg_filtered %>%
  activate(nodes) %>%
  filter(name %in% important_graph_names)
tg_main_component_check = tg_main_component_check %>% as_tibble()
tg_main_component_check = as.data.frame(tg_main_component_check)

set.seed(99)
component_1 = ggraph(tg_main_component, layout = "mds") + #mds
  geom_edge_link0(aes(width = weight), alpha = 0.3) +
  geom_node_point(aes(size = degree, fill = I(color)), shape=21, alpha=0.3) + 
  scale_edge_width(name = "Weight",range = c(0.1, 1)) +
  scale_size_continuous(name = "Degree", range = c(3, 7)) + 
  labs(title = "Largest Component Network", y="Dim 2", x = "Dim 1") +
  theme_classic() + 
     theme(axis.text.x = element_text(colour="black", size = 18)) +
     theme(axis.text.y = element_text(colour="black",size = 18)) + 
     theme(plot.title = element_text(colour="black", size=18,hjust = 0)) +
     theme(axis.title.x = element_text(colour="black", size =18)) +
     theme(axis.title.y = element_text(colour="black", size=18)) + 
    theme(legend.position = "none") 
ggsave("component_1.svg", component_1, width=4, height=4)

#Filter by institutions with at least 60 Leads
tg_directed_subgraph <- tg_filtered %>%
  activate(nodes) %>%
  filter(name %in% important_people)

tg_directed_subgraph <- tg_directed_subgraph %>%
  mutate(
    degree = centrality_degree(mode = "all"),
    degree_in = centrality_degree(mode = "in"),
    degree_out = centrality_degree(mode = "out"),
    authority = centrality_authority(weights = weight, scale = FALSE),
    eigen = centrality_eigen(weights = weight, directed=TRUE,scale = FALSE),
    pargerank = centrality_pagerank(weights = weight, directed=TRUE),
    betweenness = centrality_betweenness())

#Filter by institutions with degree > 5 due to disconnected graph and too many communities for communication 
tg_directed_subgraph <- tg_directed_subgraph %>%
  activate(nodes) %>%
  filter(degree > 5)  # Filter by degree and community

# Nodes to label
label_names <- c("National Cancer Institute (NCI)", "M.D. Anderson Cancer Center","Memorial Sloan Kettering Cancer Center","Dana-Farber Cancer Institute", "Pfizer", "Fudan University", "Sun Yat-sen University", "The Netherlands Cancer Institute", "Bayer", "Seoul National University Hospital")
label_ids <- seq_along(label_names)

# Add label id and mapping
tg_directed_subgraph <- tg_directed_subgraph %>%
  mutate(
    label_id = case_when(
      name == label_names[1] ~ "1",
      name == label_names[2] ~ "2",
      name == label_names[3] ~ "3",
      name == label_names[4] ~ "4",
      name == label_names[5] ~ "5",
      name == label_names[6] ~ "6",
      name == label_names[7] ~ "7",
      name == label_names[8] ~ "8",
      name == label_names[9] ~ "9",
      name == label_names[10] ~ "10",
      TRUE ~ NA_character_
    ),
    label_name = case_when(
      name %in% label_names ~ name,
      TRUE ~ NA_character_
    ),
    alpha_val = if_else(!is.na(label_id), 0.9, 0.7)
  )

# Create a dummy dataframe for the legend
legend_df <- tg_directed_subgraph %>%
  as_tibble() %>%
  filter(!is.na(label_id)) %>%
  distinct(label_id, label_name) %>%
  arrange(as.numeric(label_id))


weights_plot = tg_directed_subgraph %>% activate(edges) %>% pull(weight)

#Final override colours
# Your known subset of 8 communities you want to assign fixed colors to
target_communities_final <- c("7",  "25", "5",  "1",  "11", "28", "16", "6")  # Replace with your actual communities

# Generate viridis(8) palette
fixed_colors_final <- brewer.pal(10,"Paired")
fixed_colors_final <- fixed_colors_final[-c(5:6)]
fixed_colors_final = str_replace_all(fixed_colors_final,"#1F78B4","#2b94db" )

# Step 1: Create a named vector for overrides
override_colors_final <- setNames(fixed_colors_final, target_communities_final)



set.seed(234)
network_plot_publication = ggraph(tg_directed_subgraph, layout = "fr", niter=100, weight=weights_plot/10) +  # Fruchterman-Reingold layout
  geom_edge_link0(aes(width = weight), alpha = 0.4) +
  geom_node_point(aes(size = degree, fill = community, alpha = alpha_val), shape=21) +
  geom_node_text(aes(label = label_id), color="black",fontface = "bold", na.rm = TRUE) +
  scale_edge_width(name = "Weight",range = c(0.1, 2)) +
  scale_alpha_identity() +
  scale_size_continuous(name = "Degree", range = c(4, 9)) +
  scale_fill_manual(name="Community",labels = c("China","United States B", "Continental Europe", "Korea","United Kingdom", "Industry and Western\nOncology Groups","Canada","United States A"), values= override_colors_final) +
  theme_minimal() +
  labs(title = "Network Analysis of Top Lead Institutions", y="Dim 2", x = "Dim 1") +
  theme_classic() + 
     theme(axis.text.x = element_text(colour="black", size = 15)) +
     theme(axis.text.y = element_text(colour="black",size = 15)) + 
     theme(plot.title = element_text(colour="black", size=15,hjust = 0)) +
     theme(axis.title.x = element_text(colour="black", size =15)) +
     theme(axis.title.y = element_text(colour="black", size=15)) +
  annotate("text", x = -5, y = -8, hjust = 0, vjust = 0,
           label = paste0(legend_df$label_id, ": ", legend_df$label_name, collapse = "\n"),
           size = 4.5) +
  # Customize legends independently
  guides(
    fill = guide_legend(override.aes = list(size = 5)),  # larger points in fill legend
    size = guide_legend(override.aes = list(shape = 21, fill = "grey60", alpha=0.5)), 
    edge_width = guide_legend(override.aes = list(color = "grey50", size = 1))  
  ) +
  theme(legend.position = "right",
        legend.text = element_text(colour="black",size = 15),
        legend.title = element_text(colour="black",size = 15),
        legend.key.height = unit(0.5, "cm"),
        legend.key.width = unit(1, "cm"))
network_plot_publication

ggsave("network_plot_publication.svg", network_plot_publication, height = 6, width =8.5)


sub_network_data = tg_directed_subgraph %>% as.data.frame()
labels = c("China","United States B", "Continental Europe", "Korea","United Kingdom", "Industry and Western\nOncology Groups","Canada","United States A")

community_names <- data.frame(
  community = c("1",  "11", "16",  "25", "28",  "5", "6", "7"),
  community_name = c("China","United States B", "Continental Europe", "Korea","United Kingdom", "Industry and Western Oncology Groups","Canada","United States A")
)

sub_network_data <- sub_network_data %>%
  left_join(community_names, by = "community")

sub_network_data = sub_network_data %>% select(name,degree,degree_in,degree_out,eigen,community, community_name)
write.csv(sub_network_data , "sub_network_results.csv")
```
#Network plot poster
```{r}

set.seed(234)
network_plot_poster = ggraph(tg_directed_subgraph, layout = "fr", niter=100, weight=weights_plot/10) +  # Fruchterman-Reingold layout
  geom_edge_link0(aes(width = weight), alpha = 0.4) +
  geom_node_point(aes(size = degree, fill = community, alpha = alpha_val), shape=21) +
  #geom_node_text(aes(label = label_id), color="black",fontface = "bold", na.rm = TRUE) +
  scale_edge_width(name = "Weight",range = c(0.1, 2)) +
  scale_alpha_identity() +
  scale_size_continuous(name = "Degree", range = c(4, 9)) +
  scale_fill_manual(name="Community",labels = c("China","United States B", "Continental Europe", "Korea","United Kingdom", "Industry and Western\nOncology Groups","Canada","United States A"), values= override_colors_final) +
  theme_minimal() +
  labs(title = "Network Analysis of Top Lead Institutions", y="Dim 2", x = "Dim 1") +
  theme_classic() + 
     theme(axis.text.x = element_blank()) +
     theme(axis.text.y = element_blank()) + 
     theme(plot.title = element_text(colour="black", size=15,hjust = 0)) +
     theme(axis.title.x = element_blank()) +
     theme(axis.title.y = element_blank()) +
     theme(axis.ticks = element_blank()) +
     theme(axis.line = element_blank()) +
  # Customize legends independently
  guides(
    fill = guide_legend(override.aes = list(size = 5)),  # larger points in fill legend
    size = "none", 
    edge_width = "none"
  ) +
  theme(legend.position = "right",
        legend.text = element_text(colour="black",size = 15),
        legend.title = element_text(colour="black",size = 15),
        legend.key.height = unit(0.5, "cm"),
        legend.key.width = unit(1, "cm"))
network_plot_poster

ggsave("network_plot_poster.svg", network_plot_poster, height = 4, width =6.5)
ggsave("network_plot_poster.png", network_plot_poster, height = 4, width =6.5,  bg = "transparent",dpi = 600)

```
#Degree in, degree out and eigen centrality measures
```{r}
degree_plots_df_all = tg_directed %>% as_tibble()
degree_plots_df_all = as.data.frame(degree_plots_df_all)

degree_plots_df_all$name <- str_replace_all(
  degree_plots_df_all$name,
  fixed("National Cancer Institute (NCI)"),
  "NCI"
)


degree_plots_df_all$name <- str_replace_all(
  degree_plots_df_all$name,
  fixed("Eastern Cooperative Oncology Group"),
  "ECOG"
)

#Top 10 Eigenvector for all
top10_df <- degree_plots_df_all %>%
  arrange(desc(eigen)) %>%
  slice_head(n = 10)

# Reorder factor levels for plotting
top10_df$name <- factor(top10_df$name, levels = rev(top10_df$name))

eigen_all = ggplot(top10_df, aes(x = name, y = eigen)) +
  geom_segment(aes(xend = name, y = 0, yend = eigen), color = "gray70", size = 1.2) +
  geom_point(fill = top10_df$color, size = 5, shape=21) +
  geom_text(aes(label = name), y = (top10_df$eigen + 0.04), hjust = 0, size = 4.5) +
  coord_flip() +
  theme_minimal() +
  labs(x = NULL, y = "Eigenvector Centrality", title = "Top 10 Complete Network: Influence") +
  theme(axis.text.x = element_text(colour="black", size = 15)) +
     theme(axis.text.y = element_blank()) + 
     theme(plot.title = element_text(colour="black", size=15,hjust = 0)) +
     theme(axis.title.x = element_text(colour="black", size = 15)) +
     theme(axis.ticks.x = element_blank() ) +
     theme(axis.title.y = element_text(colour="black", size=15)) +
    theme(panel.grid.major.y = element_blank(),
          panel.grid.major.x = element_line(color = "grey90", size = 0.4),
          panel.grid.minor = element_line(color = "grey90", size = 0.4)) +
  scale_y_continuous(limits = c(0, max(top10_df$eigen) * 1.1), breaks = c(0,0.25,0.5,0.75,1))  # leave space for labels

ggsave("eigen_all.svg", eigen_all, height = 6, width = 4)



####Top 5 Eigenvector subgraph

degree_plots_df_sub = tg_directed_subgraph %>% as_tibble()
degree_plots_df_sub = as.data.frame(degree_plots_df_sub)

degree_plots_df_sub$name <- str_replace_all(
  degree_plots_df_sub$name,
  fixed("National Cancer Institute (NCI)"),
  "NCI"
)

top10_df <- degree_plots_df_sub %>%
  arrange(desc(eigen)) %>%
  slice_head(n = 5)

# Reorder factor levels for plotting
top10_df$name <- factor(top10_df$name, levels = rev(top10_df$name))

eigen_sub = ggplot(top10_df, aes(x = name, y = eigen)) +
  geom_segment(aes(xend = name, y = 0, yend = eigen), color = "gray70", size = 1.2) +
  geom_point(fill = top10_df$color, size = 5, shape=21) +
  geom_text(aes(label = name), y = (top10_df$eigen + 0.04), hjust = 0, size = 4.5) +
  coord_flip() +
  theme_minimal() +
  labs(x = NULL, y = "Eigenvector Centrality", title = "Top 5 of Sub-Network: Influence") +
  theme(axis.text.x = element_text(colour="black", size = 15)) +
     theme(axis.text.y = element_blank()) + 
     theme(plot.title = element_text(colour="black", size=15,hjust = 0)) +
     theme(axis.title.x = element_text(colour="black", size = 15)) +
     theme(axis.ticks.x = element_blank() ) +
     theme(axis.title.y = element_text(colour="black", size=15)) +
    theme(panel.grid.major.y = element_blank(),
          panel.grid.major.x = element_line(color = "grey90", size = 0.4),
          panel.grid.minor = element_line(color = "grey90", size = 0.4)) +
  scale_y_continuous(limits = c(0, max(top10_df$eigen) * 1.15), breaks = c(0,0.25,0.5,0.75,1)) +
  scale_x_discrete(expand = expansion(add = 0.3)) # leave space for labels

ggsave("eigen_sub.svg", eigen_sub, height = 2.75, width = 4)



####Top 5 degree subgraph

degree_plots_df_sub = tg_directed_subgraph %>% as_tibble()
degree_plots_df_sub = as.data.frame(degree_plots_df_sub)

degree_plots_df_sub$name <- str_replace_all(
  degree_plots_df_sub$name,
  fixed("National Cancer Institute (NCI)"),
  "NCI"
)

degree_plots_df_sub$name <- str_replace_all(
  degree_plots_df_sub$name,
  fixed("Memorial Sloan Kettering Cancer Center"),
  "Memorial Sloan Kettering")

top10_df <- degree_plots_df_sub %>%
  arrange(desc(degree)) %>%
  slice_head(n = 5)

# Reorder factor levels for plotting
top10_df$name <- factor(top10_df$name, levels = rev(top10_df$name))

degree_sub = ggplot(top10_df, aes(x = name, y = degree)) +
  geom_segment(aes(xend = name, y = 0, yend = degree), color = "gray70", size = 1.2) +
  geom_point(fill = top10_df$color, size = 5, shape=21) +
  geom_text(aes(label = name), y = (top10_df$degree - 0.5), hjust = 1, size = 4.5, vjust=-0.75) +
  coord_flip() +
  theme_minimal() +
  labs(x = NULL, y = "Degree Centrality", title = "Top 5 of Sub-Network: Degree Centrality") +
  theme(axis.text.x = element_text(colour="black", size = 15)) +
     theme(axis.text.y = element_blank()) + 
     theme(plot.title = element_text(colour="black", size=15,hjust = 0)) +
     theme(axis.title.x = element_text(colour="black", size = 15)) +
     theme(axis.ticks.x = element_blank() ) +
     theme(axis.title.y = element_text(colour="black", size=15)) +
    theme(panel.grid.major.y = element_blank(),
          panel.grid.major.x = element_line(color = "grey90", size = 0.4),
          panel.grid.minor = element_line(color = "grey90", size = 0.4)) +
  scale_y_continuous(limits = c(0, max(top10_df$degree) *1.1)) +
  scale_x_discrete(expand = expansion(add = 0.75)) 

ggsave("degree_sub.svg", degree_sub, height = 3, width = 4)



```

#CHORD Diagram
```{r}

# Separate out edges and node data frames
tg_nodes <-
  tg_directed_subgraph %>%
  activate(nodes) %>%
  data.frame() %>%
  rownames_to_column("rowid") %>%
  mutate(rowid = as.integer(rowid))
tg_edges <-
  tg_directed_subgraph %>%
  activate(edges) %>%
  data.frame()

named_edge_list <-
  tg_edges %>%
  # Rename from nodes
  left_join(tg_nodes, by = c("from" = "rowid")) %>%
  select(-from) %>%  # Remove unneeded column
  rename(from = name) %>%  # Rename column with names now
  
  # Rename to nodes
  left_join(tg_nodes, by = c("to" = "rowid")) %>%
  select(-to) %>%  # Remove unneeded column
  rename(to = name) %>%  # Rename column with names now

  # Cleaning up
  select(from, to, weight)

edges_chord <- tg_directed %>%
  activate(edges) %>%
  as_tibble() 
edges_chord = as.data.frame(edges_chord)

# Build square matrix from all institution pairs
all_institutions <- sort(unique(c(named_edge_list$from, named_edge_list$to)))
matrix_template <- expand.grid(from = all_institutions, to = all_institutions)

full_matrix <- matrix_template %>%
  left_join(named_edge_list, by = c("from", "to")) %>%
  mutate(weight = replace_na(weight, 0))

matrix_data <- full_matrix %>%
  pivot_wider(names_from = to, values_from = weight) %>%
  column_to_rownames("from") %>%
  as.matrix()

# Flatten the matrix into long format for coloring
long_links <- named_edge_list

# Assign link colors based on direction (lead → collaborator)
# Normalize weights to transparency scale: 0 (opaque) to 1 (transparent)
long_links <- long_links %>%
  mutate(transparency = scales::rescale(weight, to = c(0.7, 0.9)))  # adjust range as needed


# Assign base colors using from institutions
base_colors <- override_colors_final[long_links$from]
# Blend in transparency per link
link_colors <- mapply(
  FUN = adjustcolor,
  col = base_colors,
  alpha.f = long_links$transparency,
  SIMPLIFY = TRUE
)
names(link_colors) <- paste(long_links$from, long_links$to, sep = "_")

# Extract node names and their community cluster ID
community_df <- data.frame(
  name = names(membership(louvain_clusters_large)),
  community = as.factor(membership(louvain_clusters_large))
)

#Get clusters from main graph 
#Subset community from larger graph 
subset_community = community_df %>% filter(name %in% all_institutions)
subset_community$community = factor(subset_community$community)


grid_colors <- subset_community %>%
  mutate(color = override_colors_final[community]) %>%
  select(name, color) %>%
  deframe()  # returns named vector: institution → color

grid_colors = tg_directed_subgraph %>% as_tibble() %>% select(name, color) %>%
  deframe() 

#Organize nodes by cluster
# Only include institutions in your matrix
cluster_df_sorted <- subset_community %>%
  filter(name %in% rownames(matrix_data)) %>%
  arrange(community, name)

# This gives the order you want for sectors
ordered_institutions <- cluster_df_sorted$name

matrix_data_ordered <- matrix_data[ordered_institutions, ordered_institutions]
grid_col_ordered <- grid_colors[ordered_institutions]

label_names <- c("National Cancer Institute (NCI)", "M.D. Anderson Cancer Center","Memorial Sloan Kettering Cancer Center", "Pfizer", "Fudan University", "Sun Yat-sen University", "Mayo Clinic", "UNICANCER", "Bayer","Dana-Farber Cancer Institute","Massachusetts General Hospital", "Alliance for Clinical Trials in Oncology")

sector_labels <- ifelse(all_institutions %in% label_names, all_institutions, "")

split_label <- function(text, max_len = 20) {
  words <- strsplit(text, " ")[[1]]
  line <- ""
  lines <- c()
  for (word in words) {
    if (nchar(paste(line, word)) <= max_len) {
      line <- paste(line, word)
    } else {
      lines <- c(lines, trimws(line))
      line <- word
    }
  }
  lines <- c(lines, trimws(line))
  paste(lines, collapse = "\n")
}

# Clear any previous circular plots
circos.clear()

circos.par(
  "canvas.xlim" = c(-1.3, 1.3),  # default is -1 to 1
  "canvas.ylim" = c(-1.3, 1.3)   # adjust both as needed
)
circos.par(start.degree = 180)

svg(filename="chord_diagram.svg", width=16, height=16)
# Plot chord diagram
chordDiagram(
  matrix_data_ordered,
  annotationTrack = "grid",
  order = colnames(matrix_data_ordered),
  preAllocateTracks = 1,
  grid.col = grid_col_ordered,  # optional node border color
  #transparency = 0.3,
  directional = 1,      # draw directional flows
  direction.type = c("arrows"),  # add arrows only
  link.arr.type = "big.arrow", 
  link.arr.length = uh(12, "mm"), 
  link.sort = TRUE,
  link.largest.ontop = TRUE)

dev.off()

####IF YOU WANT THE TEXT
# Plot chord diagram
chordDiagram(
  matrix_data_ordered,
  annotationTrack = "grid",
  order = colnames(matrix_data_ordered),
  preAllocateTracks = 1,
  grid.col = grid_col_ordered,  # optional node border color
  #transparency = 0.3,
  directional = 1,      # draw directional flows
  direction.type = c("arrows"),  # add arrows only
  link.arr.type = "big.arrow", 
  link.arr.length = uh(12, "mm"), 
  link.sort = TRUE,
  link.largest.ontop = TRUE)

# Add labels only to selected nodes (high volume)
circos.trackPlotRegion(
  track.index = 1,
  panel.fun = function(x, y) {
    sector_name <- get.cell.meta.data("sector.index")
    if (sector_name %in% important_people) {
      label <- split_label(sector_name, max_len = 1000)  # one line break at first space
      circos.text(
        x = CELL_META$xcenter,
        y = CELL_META$ylim[1] + mm_y(2),
        labels = label,
        facing = "clockwise",
        niceFacing = TRUE,
        adj = c(0, 0.5),
        cex = 1,
        font = 1
      )
    }
  },
  bg.border = NA
)


```

#Compile common dataset for other data with no duplicate nct_id
```{r}
eligibilities = dplyr::select(eligibilities, c("nct_id", "gender", "minimum_age", "maximum_age", "healthy_volunteers"))

studies = dplyr::select(studies, c("nct_id", "study_first_submitted_date", "results_first_submitted_date", "last_update_submitted_date", "study_first_posted_date", "last_update_submitted_qc_date", "study_type", "overall_status", "phase", "enrollment", "enrollment_type", "source", "number_of_arms", "number_of_groups", "why_stopped", "has_expanded_access", "is_fda_regulated_device", "is_fda_regulated_drug", "is_unapproved_device", "is_us_export", "biospec_retention", "plan_to_share_ipd", "source_class"))

calculated_values = dplyr::select(calculated_values, c("nct_id", "number_of_facilities", "registered_in_calendar_year", "actual_duration", "were_results_reported", "months_to_report_results", "has_us_facility", "has_single_facility", "number_of_primary_outcomes_to_measure", "number_of_secondary_outcomes_to_measure"))

designs = dplyr::select(designs, c("nct_id", "allocation", "intervention_model", "observational_model", "primary_purpose", "time_perspective", "masking", "subject_masked", "caregiver_masked", "investigator_masked", "outcomes_assessor_masked"))

combo_data = merge(studies, designs, by="nct_id", all.x=TRUE, all.y=TRUE)
combo_data = merge(combo_data, eligibilities, by="nct_id", all.x=TRUE, all.y=TRUE)
combo_data = merge(combo_data, calculated_values, by="nct_id", all.x=TRUE, all.y=TRUE)

surg_onc_data = dplyr::select(ensemble_FINAL_surgonc, c("nct_id", "class", "mean"))

combo_data = merge(surg_onc_data, combo_data, by="nct_id", all.x=TRUE)

#FIX "number_of_primary_outcomes_to_measure", "number_of_secondary_outcomes_to_measure", "number_of_arms", "number_of_groups", "actua duration#####

combo_data$number_of_facilities = as.numeric(combo_data$number_of_facilities)


plot(as.factor(combo_data$class),as.numeric(combo_data$months_to_report_results))
table(combo_data$class, combo_data$allocation)

plot(as.factor(combo_data$class),as.numeric(combo_data$number_of_facilities))

```
#Tableone
```{r}
library(tableone)
library(flextable)
library(gtsummary)


## define custom test #############
fisher.test.simulate.p.values <- function(data, variable, by, ...) {
  result <- list()
  test_results <- stats::fisher.test(data[[variable]], data[[by]], simulate.p.value = TRUE)
  result$p <- test_results$p.value
  result$test <- test_results$method
  result
}
############################

 combo_data$number_of_facilities = as.numeric(combo_data$number_of_facilities)

combo_data <- combo_data %>%
  mutate(
    number_of_facilities_cut = cut(
      number_of_facilities,
      breaks = c(-Inf, 1, 5, 10, Inf),
      labels = c("Single-center", ">1–5", ">5–10", ">10"),
      right = TRUE  # Interval includes the right value (e.g., (1,5])
    )
  )

 combo_data$number_of_primary_outcomes_to_measure = as.numeric(combo_data$number_of_primary_outcomes_to_measure)
 combo_data$number_of_secondary_outcomes_to_measure = as.numeric(combo_data$number_of_secondary_outcomes_to_measure)
 combo_data$overall_status = as.factor(combo_data$overall_status)
combo_data$allocation[combo_data$allocation=="N/A"]=NA

combo_data$class = as.factor(combo_data$class)
combo_data$class = relevel(combo_data$class, "yes")

 tb3 = tbl_summary(combo_data,
             by = "class", 
             percent = "column", 
             include=c("study_type","time_perspective","allocation", "overall_status", "phase", "is_fda_regulated_device", "is_fda_regulated_drug", "biospec_retention", "plan_to_share_ipd", "were_results_reported", "has_us_facility","number_of_facilities_cut", "primary_purpose")) %>% 
   add_p(overall_status ~ "fisher.test", test.args = overall_status ~ list(simulate.p.value = TRUE)) %>%  add_n(col_label="**n/N (Missing %)**", statistic= "{N_miss} / {N_obs} ({p_miss}%)", add_footnote = "Percent") 


 tb3_flex = as_flex_table(
  tb3,
  include = everything(),
  return_calls = FALSE,
  strip_md_bold = TRUE
)

save_as_docx("surg_onc_table3" = tb3_flex,path = "/Users/danielskubleny/Documents/R projects/tensorflow_probs3/surg_onc_table_final.docx")


```


#Plot for registration per year absolute and as proportion
```{r}
table(combo_data$class, combo_data$registered_in_calendar_year)


year_proportion <- combo_data %>%
  group_by(registered_in_calendar_year, class) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(registered_in_calendar_year) %>%
  mutate(proportion = count / sum(count)) %>%
  ungroup()

# Step 4: Convert year to numeric (for line plot)
year_proportion$registered_in_calendar_year <- as.numeric(as.character(year_proportion$registered_in_calendar_year))
year_proportion$class = as.factor(year_proportion$class)
year_proportion$class = relevel(year_proportion$class, "yes")

# Step 5: Plot as line graph
proportion_year = ggplot(year_proportion, aes(x = registered_in_calendar_year, y = proportion, color = class)) +
  geom_line(size = 2) +
  geom_point(size = 2) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_color_manual(name="Study Type", label= c("Surgical Oncology", "Other"), values= c("#3B528BFF", "#27AD81FF")) +
  labs(title = "Proportion of Studies Over Time",
       x = "Year", y = "Proportion",
       color = "Outcome") +
  theme_classic() +
  theme(axis.text.x = element_text(colour="black", size = 15)) +
     theme(axis.text.y = element_text(colour="black",size = 15)) + 
     theme(plot.title = element_text(colour="black", size=15,hjust = 0)) +
     theme(axis.title.x = element_text(colour="black", size =15)) +
     theme(axis.title.y = element_text(colour="black", size=15)) +
    theme(panel.grid.major.y = element_line(color = "grey80", size = 0.4),
          panel.grid.major.x = element_line(color = "grey90", size = 0.4),
          panel.grid.minor = element_blank()) +
     theme(legend.text = element_text(colour="black", size = 14),
        legend.title = element_text(colour="black", size = 14),
        legend.background = element_rect(alpha('white', 0.4), colour="black", linewidth=0.25),
        legend.margin = margin(1,1,1,1) ,
        legend.key.height = unit(1, "lines"),
        legend.position = c(0.6,0.5)) 
ggsave("proportion_year.svg", proportion_year, height=2, width=4)


```
#Time series lead/collaborator 
```{r}
library(ggbump)
sponsors_combo = merge(combo_data,sponsors, by="nct_id")
sponsors_combo = sponsors_combo %>% filter(registered_in_calendar_year!="2024")
```
#STACKED
```{r}
# 1. Get top 10 institutions by total number of studies
top10_leads <- sponsors_combo %>%
  filter(class=="yes") %>%
  filter(lead_or_collaborator == "lead") %>%
  count(name, sort = TRUE) %>%
  top_n(10, n) %>%
  pull(name)

# 2. Prepare data: fill missing years with 0s and collapse others to "Other"
all_years <- min(sponsors_combo$registered_in_calendar_year):max(sponsors_combo$registered_in_calendar_year)

area_data <- sponsors_combo %>%
  filter(class=="yes") %>%
  filter(lead_or_collaborator == "lead") %>%
  mutate(
    registered_in_calendar_year = as.integer(registered_in_calendar_year),
    name = ifelse(name %in% top10_leads, name, "Other")
  ) %>%
  count(registered_in_calendar_year, name) %>%
  tidyr::complete(
    registered_in_calendar_year = full_seq(range(.$registered_in_calendar_year), 1),
    name,
    fill = list(n = 0)
  )

area_data = area_data %>% filter(name != "Other")

# Reorder factors so "Other" is last
area_data$name <- forcats::fct_relevel(area_data$name, top10_leads)


# 3. Plot
ggplot(area_data, aes(x = registered_in_calendar_year, y = n, fill = name)) +
  geom_area(alpha = 0.9, size = 0.3, color = "white") +
  scale_fill_brewer(palette = "Paired") +
  labs(
    title = "Number of Registered Studies by Lead Institution Over Time",
    subtitle = "Top 10 lead institutions highlighted, others grouped as 'Other'",
    x = "Year", y = "Number of Studies", fill = "Institution"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "right")
```

#FACET bump LEAD
```{r}
# Step 1: Get top 10 lead institutions overall
top10_institutions <- sponsors_combo %>%
  filter(class=="yes") %>%
  filter(lead_or_collaborator == "lead") %>%
  count(name, sort = TRUE) %>%
  slice_max(n, n = 6) %>%
  pull(name)

# Step 2: Prepare yearly rankings
rank_data <- sponsors_combo %>%
  filter(class=="yes") %>%
  filter(lead_or_collaborator == "lead", name %in% top10_institutions) %>%
  mutate(registered_in_calendar_year = as.integer(registered_in_calendar_year)) %>%  # <-- Coerce to integer
  group_by(registered_in_calendar_year, name) %>%
  summarise(studies = n(), .groups = "drop") %>%
  complete(
    registered_in_calendar_year = full_seq(range(.$registered_in_calendar_year, na.rm = TRUE), 1),
    name,
    fill = list(studies = 0)
  ) %>%
  group_by(registered_in_calendar_year) %>%
  arrange(desc(studies)) %>%
  mutate(rank = row_number()) %>%
  ungroup()

# Step 3: Expand for faceting (duplicate data for each institution)
facet_data <- expand.grid(name_highlight = top10_institutions) %>%
  left_join(rank_data, by = character()) %>%
  mutate(
    is_highlight = name == name_highlight,
    color = ifelse(is_highlight, name, "Other"),
    alpha = ifelse(is_highlight, 1, 0.2)
  )

facet_data$color <- forcats::fct_relevel(facet_data$color, "Other", after = Inf)

cols_lead = RColorBrewer::brewer.pal(9, "Blues")
cols_lead = cols_lead[4:9]

facet_data$name_highlight = str_replace_all(facet_data$name_highlight, "Memorial Sloan Kettering Cancer Center", "Memorial Sloan Kettering" )
facet_data$name_highlight = str_replace_all(facet_data$name_highlight, "M.D. Anderson Cancer Center", "M.D. Anderson" )

# Step 4: Bump plot faceted by institution
lead_overtime = ggplot(facet_data, aes(
  x = registered_in_calendar_year,
  y = studies,
  group = name,
  color = color,
  alpha = alpha
)) +
  geom_bump(size = 1.5) +
  geom_point(size = 2) +
  scale_alpha_identity() +
  scale_color_manual(values = c(cols_lead, "grey80")) +
  facet_wrap(~ name_highlight, ncol = 2) +
  labs(
    title = "Top 6 Lead Institutions Over Time",
    x = "Year", y = "Rank"
  ) +
  theme_bw(base_size = 13) +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank(),
    strip.background = element_rect(fill="#DEEBF7"),
    strip.text = element_text(colour="black", size = 16),
    axis.text.x = element_text(colour="black",size = 16),
    axis.text.y = element_text(colour="black",size = 16),
    plot.title = element_text(colour="black", size=16,hjust = 0),
    axis.title.x = element_text(colour="black", size =16),
    axis.title.y = element_text(colour="black", size=16)
  )

ggsave("lead_overtime.svg", lead_overtime, height=6, width=7.2)

```

#FACET bump Collaborator
```{r}
# Step 1: Get top 10 lead institutions overall
top10_institutions <- sponsors_combo %>%
  filter(class=="yes") %>%
  filter(lead_or_collaborator == "collaborator") %>%
  count(name, sort = TRUE) %>%
  slice_max(n, n = 6) %>%
  pull(name)

# Step 2: Prepare yearly rankings
rank_data <- sponsors_combo %>%
  filter(class=="yes") %>%
  filter(lead_or_collaborator == "collaborator", name %in% top10_institutions) %>%
  mutate(registered_in_calendar_year = as.integer(registered_in_calendar_year)) %>%  # <-- Coerce to integer
  group_by(registered_in_calendar_year, name) %>%
  summarise(studies = n(), .groups = "drop") %>%
  complete(
    registered_in_calendar_year = full_seq(range(.$registered_in_calendar_year, na.rm = TRUE), 1),
    name,
    fill = list(studies = 0)
  ) %>%
  group_by(registered_in_calendar_year) %>%
  arrange(desc(studies)) %>%
  mutate(rank = row_number()) %>%
  ungroup()

# Step 3: Expand for faceting (duplicate data for each institution)
facet_data <- expand.grid(name_highlight = top10_institutions) %>%
  left_join(rank_data, by = character()) %>%
  mutate(
    is_highlight = name == name_highlight,
    color = ifelse(is_highlight, name, "Other"),
    alpha = ifelse(is_highlight, 1, 0.2)
  )


facet_data$color <- forcats::fct_relevel(facet_data$color, "Other", after = Inf)

cols_collab = RColorBrewer::brewer.pal(9, "Greens")
cols_collab = cols_collab[4:9]

facet_data$name_highlight = str_replace_all(facet_data$name_highlight, "Memorial Sloan Kettering Cancer Center", "Memorial Sloan Kettering" )
facet_data$name_highlight = str_replace_all(facet_data$name_highlight, "M.D. Anderson Cancer Center", "M.D. Anderson" )

# Step 4: Bump plot faceted by institution
collab_overtime = ggplot(facet_data, aes(
  x = registered_in_calendar_year,
  y = studies,
  group = name,
  color = color,
  alpha = alpha
)) +
  geom_bump(size = 1.5) +
  geom_point(size = 2) +
  scale_alpha_identity() +
  scale_color_manual(values = c(cols_collab, "grey80")) +
  facet_wrap(~ name_highlight, ncol = 2) +
  labs(
    title = "Top 6 Collaborator Institutions Over Time",
    x = "Year", y = "Rank"
  ) +
  theme_bw(base_size = 13) +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank(),
    strip.background = element_rect(fill="#F7FCF5"),
    strip.text = element_text(colour="black", size = 16),
    axis.text.x = element_text(colour="black",size = 16),
    axis.text.y = element_text(colour="black",size = 16),
    plot.title = element_text(colour="black", size=16,hjust = 0),
    axis.title.x = element_text(colour="black", size =16),
    axis.title.y = element_text(colour="black", size=16)
  )

ggsave("collab_overtime.svg", collab_overtime, height=6, width=7.2)


```
#Plot proportion of countries participation 
```{r}
prop_countries_year = merge(combo_data,facilities, by="nct_id")
prop_countries_year = prop_countries_year %>% filter(class=="yes")
prop_countries_year = prop_countries_year %>% select(nct_id, class,city, state, country,registered_in_calendar_year)

```

